### 2018-06-08 paper/paper.tex
Version 0.1.

\documentclass[amsmath,superscriptaddress,showpacs,aps,prb,twocolumn]{revtex4-1}
\usepackage[colorlinks,linkcolor=blue,anchorcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{color}
\usepackage{bm}

\begin{document}
\title{Flatband ferromagnetism and dispersion relations of spin-1 excitations in topological Hubbard models}
\author{Xiao-Fei Su}
\affiliation{National Laboratory of Solid State Microstructures and Department of Physics, Nanjing University, Nanjing 210093, China}
\affiliation{School of Physics and Electronic Information, Huaibei Normal University, Huaibei 235000, China}
\author{Zhao-Long Gu}
\affiliation{National Laboratory of Solid State Microstructures and Department of Physics, Nanjing University, Nanjing 210093, China}
\author{Zhao-Yang Dong}
\affiliation{National Laboratory of Solid State Microstructures and Department of Physics, Nanjing University, Nanjing 210093, China}
\author{Shun-Li Yu}
\affiliation{National Laboratory of Solid State Microstructures and Department of Physics, Nanjing University, Nanjing 210093, China}
\affiliation{Collaborative Innovation Center of Advanced Microstructures, Nanjing University, Nanjing 210093, China}
\author{Jian-Xin Li}
\affiliation{National Laboratory of Solid State Microstructures and Department of Physics, Nanjing University, Nanjing 210093, China}
\affiliation{Collaborative Innovation Center of Advanced Microstructures, Nanjing University, Nanjing 210093, China}
\date{\today}

\begin{abstract}
\par We study the spin-1 excitation spectra of the flatband ferromagnetic phases in topological Hubbard models. As a paradigm, we consider a quarter filled square lattice Hubbard model whose free part is the $\pi$ flux model with topologically nontrivial and nearly-flat electron bands. The model Hamiltonian either explicitly breaks the time-reversal symmetry but preserves the spin $SU(2)$ rotation symmetry (Chern Hubbard model), or preserves the time-reversal symmetry but explicitly breaks the spin $SU(2)$ rotation symmetry ($Z_2$ Hubbard model). By using the numerical exact diagonalization method with a projection onto the lower nearly-flat electron band, we determine the critical Hubbard interaction strength upon which the ferromagnetic phase is stable, and elaborate the ferromagnetic spin-1 excitation spectra of the Chern Hubbard model and the $Z_2$ Hubbard model. Both spectra consist of collective modes (spin waves) and individual modes (Stoner continuum). For the Chern Hubbard model, the spin wave is gapless while for the $Z_2$ Hubbard model, the spin wave is gapped. Remarkably, for both cases, the nonflatness of the free electron bands introduces dips of the lower boundary of the Stoner continuum, \textcolor{red}{and the scatterings between the Stoner continuum at these dips and the collective modes significantly reduce the latter's energy}, which can lead to roton-like spin wave excitations. With the increase of this nonflatness, the energy of the induced roton-like modes goes down and finally touches zero, which results in the destabilization of the ferromagnetic phase.
\end{abstract}
\maketitle

\section{Introduction}
\par Electronic bands with nonzero topological indices reside on the center of a substantial amount of topological phenomena in condensed matter physics \cite{HK_RMP2010,QZ_RMP2011}. It was proposed in a pioneering work by Haldane \cite{H_PRL1988} that a spinless fermionic model on a honeycomb lattice exhibits integer quantum Hall effect \cite{KDP_PRL1980} without an external magnetic field. This model, serving as the first example of Chern insulator, breaks the time reversal symmetry with a complex next nearest neighbor hopping and is characterized by a nonzero Chern number \cite{TKNN_PRL1982}. Later, the concept was generalized to time reversal symmetric systems with spin-orbital coupling (SOC), such as the monolayer graphene \cite{KM_PRL2005a,KM_PRL2005b} and HgTe/CdTe quantum wells \cite{BHZ_S2006,Ketc_S2007}. The SOC there generates complex hopping terms similar to that proposed by Haldane but with opposite chiralities for electrons with up spins and down spins, resulting in the quantum spin Hall insulator characterized by a $Z_2$ index.

\par The lattice models with nontrivial band topology share much similarity with the two-dimensional electron gas (2DEG) under a strong magnetic field with Landau levels, e.g. the existence of topologically protected gapless edge states \cite{H_PRL1988,L_PRB1981,H_PRB1982,KM_PRL2005a,KM_PRL2005b,BHZ_S2006,Ketc_S2007}. Thus more novel phases other than the Chern insulator or $Z_2$ insulator are expected when Coulomb interactions are taken into account, as is similar to the fractional quantum Hall effect \cite{TSG_PRL1982,L_PRL1983} in the 2DEG with Landau levels. However, different from Landau levels, energy bands in lattice models usually have noneligible dispersions, which weakens the effect of Coulomb interactions. Therefore, in recent years, much effort has been devoted to the design and search of tight-binding models that host nearly-flat electron bands with nontrivial topology \cite{TMW_PRL2011,WR_PRB2011,SGKD_PRL2011,WGGS_PRL2011,NSCM_PRL2011,TB_PRB2012,YGSD_PRB2012}. Analogous exotic phases, such as the fractional Chern insulator and fractional topological insulator were numerically verified to emerge in such nearly-flat topological bands when strong Coulomb interactions are turned on \cite{NSCM_PRL2011,SGSS_NC2011,RB_PRX2011,NSRCM_PRB2011,LBFL_PRL2012}.

\par Another involved intriguing phenomenon arising from Coulomb repulsions in flat or nearly-flat bands is the itinerant ferromagnetism \cite{T_PTP1998}. It was proved by Tasaki and Mielke that the ground state of an flat electron band with its filling factor not more than but sufficiently close to $1/2$ is ferromagnetically ordered as long as an infinitesimal onsite Hubbard repulsion is present \cite{T_PRL1992,M_PLA1993,MT_CMP1993}. Afterwards this ferromagnetism was shown to be stable against small nonflatness of the electron bands if and only if the Hubbard interaction exceeds a critical value \cite{T_PRL1994}. Spin wave excitations over this ferromagnetic ground state was also studied \cite{KA_PRL1994,SGDL_arXiv2008} and itinerant topological magnons has been reported quite recently \cite{SGDL_arXiv2008}.

\par The interplay between flatband ferromagnetism and nontrivial band topology enriches the related physics. In fact, ferromagnetism is essential in the generation of stable fractional Chern insulators in the proposals where the spin degrees of freedom of electrons are considered \cite{TMW_PRL2011,NSRCM_PRB2011,LBFL_PRL2012}. Furthermore, ferromagnetism can also lead to possible high-temperature quantum anomalous Hall effect (QAHE) when the nearly-flat topological band is half-filled \cite{NSRCM_PRL2012}. In this paper, we focus on this ferromagnetism induced QAHE in nearly-flat topological bands. As a paradigm, we consider a square lattice Hubbard model whose free part is the $\pi$ flux model with topologically nontrivial and nearly-flat electron bands. The model Hamiltonian either explicitly breaks the time-reversal symmetry but preserves the spin $SU(2)$ rotation symmetry (Chern Hubbard model), or preserves the time-reversal symmetry but explicitly breaks the spin $SU(2)$ rotation symmetry ($Z_2$ Hubbard model). When the model is quarter filled (or correspondingly, the lower nearly-flat band is half filled), the ground state is spin fully polarized due to the ferromagnetism and exhibits QAHE because of the nonzero Chern number of a single-spin band. Then the low energy physics is dominated by the one-spin-flip excitations. These spin-1 excitations has been studied by a generalized bosonization scheme where the interacting fermionic model is mapped to a free bosonic model describing spin-wave excitations at the harmonic approximation \cite{DG_PRB2015}. The ferromagnetism was shown to be stable against such spin wave excitations, which are gapless in the Chern Hubbard model and gapped in the $Z_2$ Hubbard model. However in this bosonization scheme, the free part of the electron model plays no role in the spin wave excitations other than contributing a global constant, which means it should fail due to the competition between the kinetic energy and potential energy of electrons when the nonflatness of the electron bands is not negligible \cite{T_PRL1994}. What's worse, in a strictly local periodic tight-binding model, an energy band with a nonzero Chern number cannot be exactly flat \cite{CMST_JPMT2014}. Therefore it remains an open question on whether the ground state is stable against the spin-1 excitations and how the the nonflatness of the electron bands manifests itself in the spin-1 excitation spectra in such models.

\par To elucidate these questions, we adopt the numerical exact diagonalization method with a projection onto the lower nearly-flat band to take close investigations on the spin-1 excitations of the models. A critical Hubbard interaction strength is found for both the Chern Hubbard model and the $Z_2$ Hubbard model, upon which the ferromagnetic phase is stable. Furthermore the spin-1 excitation spectra are shown to consist of collective modes (spin waves) and individual modes (Stoner continuum). For the Chern Hubbard model, the spin wave is gapless while for the $Z_2$ Hubbard model, the spin wave is gapped. Remarkably, for both cases, the nonflatness of the free electron bands introduces dips of the lower boundary of the Stoner continuum, \textcolor{red}{and the scatterings between the Stoner continuum at these dips and the collective modes significantly reduce the latter's energy}, which can lead to roton-like spin wave excitations. With the increase of this nonflatness, the energy of the induced roton-like modes goes down and finally touches zero, which results in the destabilization of the ferromagnetic phase.

\par The rest of the paper is organized as follows. In Sec. \ref{mm}, we introduce the Chern Hubbard model and $Z_2$ Hubbard model studied in this paper, discuss the nontrivial band topology of their free parts, interpret the emergence of QAHE resulting from the interplay of flat-band ferromagnetism and nontrivial band topology, and formulate the exact diagonalization method with a projection onto the lower nearly-flat band on details. In Sec. \ref{nr}, we summarize the phase diagram and elaborate the spin-1 excitation spectra of both models. Section \ref{sd} provides a summary and discussion.

\section{Model and method}\label{mm}
\subsection{Introduction to model}\label{intro_to_model}

\begin{figure}
\includegraphics[width=\columnwidth]{lattice}
\caption{(Color online) (a) Schematic representation of the $\pi$-flux model. Blue and red solid circles denote the $A$ and $B$ sublattices, respectively. The nearest-neighbor hopping amplitudes (solid black lines) are equal to $t_{1}\exp[i\alpha(\sigma)\pi/4]$ ($\alpha(\sigma)=1$ for the time-reversal-symmetry-breaking case and $\alpha(\sigma)=\pm1$ for the time-reversal-symmetry-preserving case) along the direction of the arrows, the next-nearest-neighbor hopping amplitudes are equal to $t_{2}$ (dashed green lines) and $-t_{2}$ (dashed purple lines). The shaded area denotes the unit cell. (b) First Brillouin Zone. $\Gamma=(0,0)$, $X_{1,2}=(\pm\pi,0)$, $Y_{1,2}=(0,\pm\pi)$, $M_{1,2,3,4}=(\pm\pi,\pm\pi)$.}
\label{lattice}
\end{figure}

\par We consider a generalized $\pi$-flux Hubbard model on the square lattice, whose Hamiltonian can be written as $H=H_0+H_U$, where $H_0$ is the generalization with electron spin of the original spinless model proposed in Ref. \cite{NSCM_PRL2011},
\begin{equation}
H_0=\sum_{\langle ij\rangle,\sigma}(t_1^{ij,\sigma}c_{i\sigma}^\dagger c_{j\sigma}+\text{H.c.})
    +\sum_{\langle\langle ij\rangle\rangle,\sigma}(t_2^{ij}c_{i\sigma}^\dagger c_{j\sigma}+\text{H.c.}),
\end{equation}
and $H_U$ is the Hubbard interaction
\begin{equation}
H_U=U\sum_{i}n_{i\uparrow}n_{i\downarrow}.
\end{equation}
Here, $c_{i\sigma}^{\dagger}(c_{i\sigma})$ creates (annihilates) a spin $\sigma$ electron at site $i$, $n_{i\sigma}$ is the spin $\sigma$ electron number operator at site $i$, $\langle ij\rangle$ denotes the nearest-neighbor (NN) bonds and $\langle\langle ij\rangle\rangle$ denotes the next-nearest-neighbor (NNN) bonds. As is shown in Fig. \ref{lattice}(a), the spin-dependent NN hopping amplitude $t_1^{ij,\sigma}$ and the spin-independent NNN hopping amplitude $t_2^{ij}$ are given by
\begin{equation}
t_1^{ij,\sigma}=t_1\exp\left[i(-1)^{\delta^1_{ij}}\alpha_\sigma\pi/4\right],
\end{equation}
\begin{equation}
t_2^{ij}=t_2(-1)^{\delta^2_{ij}}.
\end{equation}
Here, $\delta^1_{ij}=+1$ if the NN electron hopping is along the direction of the solid black arrow and $\delta^1_{ij}=-1$ if along the reversed direction. $\delta^2_{ij}=+1$ if the NNN electron hopping is along the dashed green lines and $\delta^2_{ij}=-1$ if along the dashed purple lines. The spin-dependent phase $\alpha_\sigma$ breaks the time-reversal symmetry but preserves the spin $SU(2)$ rotation symmetry if $\alpha_\uparrow=\alpha_\downarrow=+1$, whereas it preserves the time-reversal symmetry but breaks the spin $SU(2)$ rotation symmetry if $\alpha_\uparrow=+1$ and $\alpha_\downarrow=-1$.

\par Due to the complex NN hopping, each electron will acquire a $\pi$ phase as it hops around a plaquette along the direction of the black arrows as indicated in Fig. \ref{lattice}(a). Therefore, $H_0$ describes free electrons hopping on a square lattice in the presence of a fictitious staggered $\pi$-flux pattern \cite{WWZ_PRB1989}. For the time-reversal-symmetry-breaking case, $\alpha_\uparrow=\alpha_\downarrow$, the flux experienced by spin-up electrons and spin-down electrons is the same, while for the time-reversal-symmetry-preserving case, $\alpha_\uparrow=-\alpha_\downarrow$, it is opposite.

\subsection{Topology of free part}\label{topo_free_part}
\par Gapped noninteracting fermionic systems can be topologically classified in the presence of symmetries by their Hamiltonians in the momentum space\cite{SRFL_PRB2008}. After the Fourier transformation, the free part $H_0$ of our model reads
\begin{equation}
H_0=\sum_{\mathbf{k}\sigma}\psi^{\dagger}_{\mathbf{k}\sigma}h_{\mathbf{k}\sigma}\psi_{\mathbf{k}\sigma},
\end{equation}
where $\psi^{\dagger}_{\mathbf{k}\sigma}=(c^{\dagger}_{A\mathbf{k}\sigma},c^{\dagger}_{B\mathbf{k}\sigma})$ and
\begin{equation}
h_{\mathbf{k}\sigma}={\mathbf{D}_{\mathbf{k}\sigma}}\cdot\bm{\tau}.
\end{equation}
Here $\bm{\tau}=(\tau_{1},\tau_{2},\tau_{3})$ is a $2\times2$ matrix vector, $\tau_{1},\tau_{2},\tau_{3}$ are the three Pauli matrices for the sublattice degrees of freedom. The components of $\mathbf{D}_{\mathbf{k}\sigma}$ are given by
\begin{equation}
\begin{aligned}
&D_{1,\mathbf{k}}=2\sqrt{2}t_{1}\cos\frac{k_{x}}{2}\cos\frac{k_{y}}{2},\\
&D_{2,\mathbf{k}}=2\sqrt{2}t_{1}\alpha_\sigma\sin\frac{k_{x}}{2}\sin\frac{k_{y}}{2},\\
&D_{3,\mathbf{k}}=2t_{2}(\cos k_{x}-\cos k_{y}).
\end{aligned}
\end{equation}
$H_0$ can be diagonalized with the transformation
\begin{equation}
\begin{aligned}
&c_{A\mathbf{k}\sigma}=\mu_{1,\mathbf{k}\sigma}d_{\mathbf{k}\sigma}+\mu_{2,\mathbf{k}\sigma}f_{\mathbf{k}\sigma},\\
&c_{B\mathbf{k}\sigma}=\mu^{\ast}_{1,\mathbf{k}\sigma}f_{\mathbf{k}\sigma}-\mu^{\ast}_{2,\mathbf{k}\sigma}d_{\mathbf{k}\sigma},\\
\end{aligned}
\end{equation}
where
\begin{equation}
\begin{aligned}
&\mu_{1,\mathbf{k}\sigma}=\frac{D_{1,\mathbf{k}}-i\alpha_\sigma D_{2,\mathbf{k}}}{\sqrt{2D_{\mathbf{k}}(D_{\mathbf{k}}+D_{3,\mathbf{k}})}},\\
&\mu_{2,\mathbf{k}\sigma}=\frac{D_{\mathbf{k}}+D_{3,\mathbf{k}}}{\sqrt{2D_{\mathbf{k}}(D_{\mathbf{k}}+D_{3,\mathbf{k}})}},\\
\end{aligned}
\end{equation}
with $D_{\mathbf{k}}=\sqrt{D^2_{1,\mathbf{k}}+D^2_{2,\mathbf{k}}+D^2_{3,\mathbf{k}}}$. The diagonalized $H_0$ is given by
\begin{equation}
H_{0}=\sum_{\mathbf{k}\sigma}\varepsilon_{d}(\mathbf{k})c^{\dagger}_{\mathbf{k}\sigma}c_{\mathbf{k}\sigma}
      +\sum_{\mathbf{k}\sigma}\varepsilon_{f}(\mathbf{k})f^{\dagger}_{\mathbf{k}\sigma}f_{\mathbf{k}\sigma},
\end{equation}
where $\varepsilon_{d}(\mathbf{k})=-D_{\mathbf{k}}$, $\varepsilon_{f}(\mathbf{k})=D_{\mathbf{k}}$. It can be seen that there exists a gap between the $d$ band and $f$ band when $t_1\ne0$ and $t_2\ne0$.
\par When there is a gap between the $d$ band and the $f$ band, these bands can be shown to be topologically non-trivial by calculating their Chern numbers \cite{TKNN_PRL1982} (for the time-reversal-symmetry breaking case) or $Z_2$ indices \cite{KM_PRL2005b,SWSH_PRL2006} (for the time-reversal-symmetry preserving case). The Chern number for a single spin component of the $d$ band or the $f$ band can be expressed in terms of the coefficients $D_{i,\mathbf{k}}$ \cite{HK_RMP2010,QZ_RMP2011},
\begin{equation}
C_{\sigma}^{d/f}=\pm\frac{1}{4\pi}\int_{BZ}d^{2}k\;
    \hat{\mathbf{D}}_{\mathbf{k}\sigma}\cdot(\partial_{k_{x}}\hat{\mathbf{D}}_{\mathbf{k}\sigma}\times\partial_{k_{y}}\hat{\mathbf{D}}_{\mathbf{k}\sigma})
    =\pm\alpha_\sigma,
\end{equation}
with $\hat{\mathbf{D}}_{\mathbf{k}\sigma}\equiv\mathbf{D}_{\mathbf{k}\sigma}/D_{\mathbf{k}}$.
\par When the system breaks the time-reversal symmetry, i.e. $\alpha_\uparrow=\alpha_\downarrow=1$, $C_\uparrow^d=C_\downarrow^d=1$ and $C_\uparrow^f=C_\downarrow^f=-1$, the total Chern number of the $d$ band is $C^d=C_\uparrow^d+C_\downarrow^d=2$. Therefore, the ground state of $H_0$ will be a noninteracting Chern insulator and exhibits quantum anomalous Hall effect(QAHE) when the lower $d$ band is fully filled. When the system preserves the time-reversal symmetry, i.e. $\alpha_\uparrow=-\alpha_\downarrow=1$, $C_\uparrow^d=-C_\downarrow^d=1$ and $C_\uparrow^f=-C_\downarrow^f=-1$, the total Chern number of the $d$ band is zero. However, the $Z_2$ index, which is defined as
\begin{equation}
\nu=\frac{1}{2}(C_{\uparrow}-C{\downarrow})\mod2,
\end{equation}
of the $d$ band is $1$ and nontrivial. As a consequence, the the ground state of $H_0$ will be a noninteracting $Z_2$ insulator and exhibits quantum spin Hall effect(QSHE) when the lower $d$ band is fully filled.

\subsection{Emergence of QAHE in half-filled nearly-flat topological bands}\label{emerg_qahe}
\par For a free fermionic system hosting an energy band with a nonzero Chern number or $Z_2$ index, the distinguished phenomenon resulting from this nontrivial band topology, such as QAHE or QSHE, only manifests itself when the topological band is fully filled. At any fractional filling, the ground state of such a system will be a trivial metal. Intriguingly, when the Coulomb interactions between electrons step in, the physics of the nontrivial band topology becomes more involved, especially when the band is nearly flat so that the effects of the Coulomb interactions are highly enhanced. Combined with strong Coulomb interactions, nontrivial topological phases can emerge from fractionally filled topological bands. In this article, we are interested in half-filled strongly-correlated nearly-flat topological bands where QAHE can arise\cite{NSCM_PRL2011,DG_PRB2015}. The essence for the occurrence of this nontrivial phase is the emergence of itinerant ferromagnetism on nearly-flat bands\cite{T_PTP1998}, which fully polarizes all the electron spins. Therefore only one spin component of the topological band will be fully filled exactly, which leads to QAHE due to the nonzero Chern number of that spin component of the band.

\par The Chern Hubbard model and $Z_2$ Hubbard model on square lattice described above serve as the paradigm, where half filling of the lower electron band corresponds to quarter filling of the whole system because of the existence of $AB$ sublattices. When $t_2/t_1$ takes values in a selected region, the $d$ band and $f$ band are quite flat in that the flatness ratio $\Delta/W$, which is defined as the ratio between the electron gap $\Delta$ between these two bands and the bandwidth $W$ of the lower band, can be as large as $4.83$ \cite{NSCM_PRL2011,DG_PRB2015}. In Ref. \cite{NSCM_PRL2011}, the quarter filled $Z_2$ Hubbard model was studied and all spin excitations were shown to be gapped in the flatband limit, which is essential for a possible high-temperature realization of QAHE. In Ref. \cite{DG_PRB2015}, both models at quarter filling were considered and a generalized bosonization scheme was developed to study the spin-1 excitation spectra. However, in their bosonization scheme, the free part $H_0$ of the Hamiltonian only contributes a constant to the spectra, thus it cannot capture the physics caused by the nonflatness of the topological electron bands, which is unavoidable because an energy band with a nonzero Chern number in a strictly local periodic tight binding model cannot be exactly flat \cite{CMST_JPMT2014}. The spin-1 excitations, as the dominant low-energy excitations, deserves closer investigations to understand the stability of such phases and the physics related to the nonflatness of the topological bands. In the next subsection, we will introduce the exact diagonalization method with a projection onto the lower nearly-flat electron band to elucidate these questions.

\subsection{Exact diagonalization with projection}\label{exact_diag_proj}
\par Exact diagonalization method with a projection onto the low-energy Hilbert space bas been widely applied to systems that host flat or nearly-flat energy bands \cite{NSCM_PRL2011,RB_PRX2011,NSRCM_PRB2011,SGDL_arXiv2008,NSRCM_PRL2012,DG_PRB2015}. This approach ignores the band mixing of electrons and applies when the Coulomb interaction is small compared to the energy gap of the electron bands.

\par For the model we study in this article, the relevant low-energy subspace is the lower electron band, i.e. the $d$ band. Let $P$ denote the corresponding projector, then the Hamiltonian after the projection is
\begin{widetext}
\begin{equation}
P^\dagger HP=\sum_{\mathbf{k}\sigma}\varepsilon_d(\mathbf{k})d^{\dagger}_{\mathbf{k}\sigma}d_{\mathbf{k}\sigma}
+\frac{U}{N}\sum_{a=1,2}\sum_{\mathbf{k}\mathbf{k}^{\prime}\mathbf{q}}
(\mu^{\ast}_{a,\mathbf{k}+\mathbf{q}\uparrow}\mu^{\ast}_{a,\mathbf{k}^{\prime}-\mathbf{q}\downarrow}\mu_{a,\mathbf{k}^{\prime}\downarrow}\mu_{a,\mathbf{k}\uparrow})
d^{\dagger}_{\mathbf{k}+\mathbf{q}\uparrow}d^{\dagger}_{\mathbf{k}^{\prime}-\mathbf{q}\downarrow}d_{\mathbf{k}^{\prime}\downarrow}d_{\mathbf{k}\uparrow}.
\end{equation}
\end{widetext}
Let $|\text{FM}\rangle$ denote the spin-up fully polarized state on the $d$ band,
\begin{equation}
|\text{FM}\rangle=\prod_{\mathbf{k}\in \text{FBZ}}d^{\dagger}_{\mathbf{k}\uparrow}|\text{0}\rangle,
\end{equation}
where FBZ denotes the first Brillouin zone [see Fig. \ref{lattice}(b)] and $|0\rangle$ is the fermion vaccum. Then the basis of the spin-1 excitations with a center-of-mass momentum $\mathbf{q}$ over this reference state can be written as
\begin{equation}
|\mathbf{k}_i\rangle_\mathbf{q}=d^{\dagger}_{\mathbf{k}_{i}-\mathbf{q}\downarrow}d_{\mathbf{k}_{i}\uparrow}|\text{FM}\rangle.
\end{equation}
Here $d^{\dagger}_{\mathbf{k}_{i}-\mathbf{q}\downarrow}$ creates a spin-down electron with momentum $\mathbf{k}_{i}-\mathbf{q}$ and $d_{\mathbf{k}_{i}\uparrow}$ annihilates a spin-up electron with momentum $\mathbf{k}_{i}$. Therefore this basis labels a spin-1 scattering channel with the index $\mathbf{k}_{i}$. Thus the dimension of this Hilbert space scales linearly with respect to the number of electron momentums \cite{SGDL_arXiv2008}, and a much larger system can be numerically accessed than the usual exact diagonalization without projection, which enables us to analyze the properties of the spin-1 excitation spectra at the whole first Brillouin zone rather than some restricted discrete points solely. The matrix element of the projected Hamiltonian on this spin-1 excitation basis can be easily obtained after some algebra,
\begin{widetext}
\begin{equation}\label{PHP}
\begin{aligned}
_\mathbf{q}\langle\mathbf{k}_j|P^\dagger HP|\mathbf{k}_i\rangle_\mathbf{q}&=\left[\varepsilon_d(\mathbf{k}_i-\mathbf{q})-\varepsilon_d(\mathbf{k}_i)
+\frac{U}{N}\sum_{a=1,2}\sum_{\mathbf{p}\neq\mathbf{k}_i}\left|\mu_{a,\mathbf{p}\uparrow}\right|^2\left|\mu_{a,\mathbf{k}_{i}-\mathbf{q}\downarrow}\right|^2
\right]\delta_{\mathbf{k}_j,\mathbf{k}_i} \\
& -\frac{U}{N}\sum_{a=1,2}
\mu^{\ast}_{a,\mathbf{k}_{i}\uparrow}\mu_{a,\mathbf{k}_{j}\uparrow}\mu^{\ast}_{a,\mathbf{k}_j-\mathbf{q}\downarrow}\mu_{a,\mathbf{k}_i-\mathbf{q}\downarrow}
\left(1-\delta_{\mathbf{k}_j,\mathbf{k}_i}\right).
\end{aligned}
\end{equation}
\end{widetext}
Here $\delta_{\mathbf{k}_j,\mathbf{k}_i}$ is the Kronecker delta function. Then the full spin-1 excitation spectra can be obtained by the diagonalization of the matrix whose elements are defined by Eq. (\ref{PHP}). It is noted that $|\text{FM}\rangle$ is the true ground state only if the whole spin-1 excitation spectra have no negative eigen energies. Thus we can use this as the criterion to determine the destabilization point of the ferromagnetic phase.
\par We also want to give some remarks on the flatband limit in the framework of this method. The free part $h^\text{flat}_{\mathbf{k}\sigma}$ of the Hamiltonian in the flatband limit is defined as
\begin{equation}
h^{\text{flat}}_{\mathbf{k}\sigma}=\frac{h_{\mathbf{k}\sigma}}{\left|\varepsilon_{d}(\mathbf{k})\right|}={\hat{\mathbf{D}}_{\mathbf{k}\sigma}}\cdot\mathbf{\tau}.
\end{equation}
$h^\text{flat}_{\mathbf{k}\sigma}$ shares the same eigenfunctions with $h_{\mathbf{k}\sigma}$ but has exactly-flat energy bands with the eigenvalues being $\pm 1$. $h^\text{flat}_{\mathbf{k}\sigma}$, together with the Hubbard interaction $H_U$, defines the flatband limit of the original Hamiltonian. To approach this limit, long-range hopping terms in the real space must be included \cite{NSCM_PRL2011}. To elaborate the physics related to the nonflatness of the $d$ band, we also calculated the spin-1 excitation spectra in the flatband limit for comparison. This can be done by simply ignoring the $\left[\varepsilon_d(\mathbf{k}_i-\mathbf{q})-\varepsilon_d(\mathbf{k}_i)\right]\delta_{\mathbf{k}_j,\mathbf{k}_i}$ term in Eq. (\ref{PHP}), because of the same eigenfunctions shared by $h^\text{flat}_{\mathbf{k}\sigma}$ and $h_{\mathbf{k}\sigma}$.

\section{Numerical results}\label{nr}
\begin{figure}
\includegraphics[width=\columnwidth]{phase}
\caption{(Color online) Phase diagram of the quarter-filled (a) Chern Hubbard model and (b) $Z_2$ Hubbard model. The colormap represents the ratio $U/W$ of the lower electron band, where $U$ is the Hubbard interaction strength and $W$ is the lower electron bandwidth. Red stars in (a) and (b) mark the parameters used in Fig. \ref{fmcispectrum} and Fig. \ref{tispectrum}(a), respectively. Blue stars in (a) and (b) mark the parameters used in Fig. \ref{pbcispectrum}. and Fig. \ref{tispectrum}(b), respectively.}
\label{phase}
\end{figure}

\par The phase diagrams of the quarter-filled Chern Hubbard model and $Z_2$ Hubbard model are shown in Fig. \ref{phase}. Here, the colormap represents the ratio of the Hubbard interaction strength $U$ over the lower electron bandwidth $W$. FM denotes the ferromagnetic phase and NFM denotes the non-ferromagnetic phase. As is discussed in Sec. \ref{exact_diag_proj}, the phase boundary is determined as the onset point of $U$ below which negative spin-1 excitation spectra appear. It is clear that a critical Hubbard interaction strength is needed to maintain the ferromagnetically ordered ground state when the electron band has finite nonflatness. This observation is expected as a result from the competition between the kinetic energy and the potential energy of the electrons, as a fully spin-polarized state minimizes the energy of Hubbard interactions but cost more energy when the electron band disperses. An intriguing fact about the phase diagrams is that the phase boundaries always lie near the contour line with $U/W=2.0$.

\subsection{Chern Hubbard model}
\begin{figure}
\includegraphics[width=\columnwidth]{fmcispectrum}
\caption{(Color online) (a) Spin-1 excitation spectra of the Chern Hubbard model in the FM phase. Green solid lines denote the spin waves and the grey regions denote the Stoner continuum. Green dashed lines represent the corresponding spectra in the flatband limit. Blue solid lines denote the upper and lower boundaries of the Stoner continuum determined by the split band picture shown in (c). Black arrows mark the local minima of the Stoner continuum. (b) Spectral weights for the lowest four eigen levels of the spin-1 excitation spectra with the center-of-mass momentum $\mathbf{q}=(0,0)$. (c) Illustration of the splitting of the lower electron bands for up spins (blue solid line) and down spins (blue dashed line) in the presence of ferromagnetism of the Chern Hubbard model. Inset shows the positions of the corresponding local maxima (red solid circles) of the spin-up bands and local minima (purple solid circles) of the spin-down bands in the first Brillouin zone. Green arrows denote the corresponding scattering channels of the dips of the Stoner continuum marked in (a). (d) Differences of the spectral weights subtracted by those in the flatband limit for the lowest four eigen levels of the spin-1 excitation spectra with the center-of-mass momentum $\mathbf{q}=(\pi,0)$. The parameters are $t_1=1.0$, $t_2=0.49$ and $U=2.3$.}
\label{fmcispectrum}
\end{figure}

\par To gain a comprehensive understanding of the physics observed above, the spin-1 excitation spectra along a high symmetry path in the first Brillouin zone of the quarter-filled Chern Hubbard model with two different $t_2$ values are plotted in Fig. \ref{cispectrum}(a)-(b) as the green solid lines and the shaded areas. For a comparison, the spectra for those in the corresponding flat-band limit are shown as the green dashed lines as well. Apparently, for both the original model and the flat-band limit model, the spectra consist of two types of modes: the low-lying ones exhibiting well-defined band structures and the high-energy ones forming a continuum. The former are identified as the magnons with an acoustic branch and an optical branch while the latter are identified as the Stoner continuum. At $\Gamma$ point, the acoustic magnon band is gapless, which is the character of a ferromagnetic excitation with the Goldstone mode as the result of the spontaneous spin $SU(2)$ symmetry breaking. Remarkably,

\begin{figure}
\includegraphics[width=\columnwidth]{pbcispectrum}
\caption{(Color online) (a) Spin-1 excitation spectra of the Chern Hubbard model at the FM/NFM phase boundary. Black arrow marks the minimum of the Stoner continuum. Inset shows its low-energy part along $\Gamma$-$X$ path with an amplified resolution. (b) Illustration of the splitting of the lower electron bands of the Chern Hubbard model. Inset shows the positions of the corresponding local maxima (red solid circles) of the spin-up bands and local minima (purple solid circles) of the spin-down bands in the first Brillouin zone. Green arrows denote the corresponding scattering channels of the minimum of the Stoner continuum marked in (a). The parameters are $t_1=1.0$, $t_2=0.43$ and $U=2.3$.}
\label{pbcispectrum}
\end{figure}

%\par The Hamiltonian has SU(2) spin rotation symmetry. The ferromagnetic ground state breaks SU(2) spin rotation symmetry to U(1) symmetry, and one would therefore expect the presence of a Goldstone mode in the spin-wave excitation spectrum due to spontaneous broken symmetry. We show such a mode at the center of the Brillouin Zone [see Fig.~\ref{fig3}]. The lowest two bands (black solid line and red dashed line) respectively corresponds to the acoustic and optical magnon (spin-wave excitation) bands degenerating along the $X-M$ path in the Brillouin Zone while the higher excitations represent to Stoner continuum in Fig.~\ref{fig3}. $G_{1}(\mathbf{k}_{i},\mathbf{Q})$ dominates the Stoner excitation gap while $G_{2}(\mathbf{k}_{i},\mathbf{k}_{j},\mathbf{Q})$ has very little contribution to the Stoner gap. With the increase of $W/U$, the optical spin-wave excitation is eventually absorbed in the Stoner continuum along the $\Gamma-X$ path in the Brillouin Zone [see Fig. \ref{fig3}(a),(b),(c)]. For $U_{c}=2t_{1}$ ($W/U_{c}=0.414$), there is a critical situation of ferromagnetic ground state, consistent with the theory of Tasaki \cite{Tasaki1} [see Fig.~\ref{fig3}(c)]. When $W/U>W/U_{c}$, the spin-wave excitation appears to be negative somewhere, and the ground state is no longer fully spin-polarized ferromagnetic state.  In order to investigate the effect of the $c$ band to the spin-1 excitation spectrum, we have to consider the case that the $c$ band is assumed as a completely flat-band. In this case, the excitation energies are proportional to the on-site repulsion $U$ instead of the nearest-neighbor hopping energy $t_{1}$ and the boundary of Stoner continuum coalesces into a line [see Fig.~\ref{fig3}(d)]. No matter how small the repulsion $U$ is, the ground state is a fully spin-polarized ferromagnetic state. In contrast to this case, the dispersion of the $c$ band makes the excitation spectrum have the downwarping behavior and gives rise to the phase transition for $U=2t_{1}$ [see Fig.~\ref{fig3}(a),(b),(c)]. The spin-wave excitation spectrum in Fig.~\ref{fig3}(d) is qualitatively similar to the result of the bosonization formalism in ref.~\cite{Doretto}. However, Our result takes into account the effect of the spin-wave interaction in the bosonization formalism. The bosonization formalism neglects the contribution of the nearly flat-band to the spin wave excitation and can't determine what the $W/U$ value is, the ground state is the fully spin-polarized ferromagnetic state.

\subsection{Topological Hubbard model}
\begin{figure}
\includegraphics[width=\columnwidth]{tispectrum}
\caption{(Color online) (a) and (b): Spin-1 excitation spectra of the $Z_2$ Hubbard model. Green solid lines denote the spin waves and the grey regions denote the Stoner continuum. Green dashed lines represent the corresponding spectra in the flatband limit. Black arrows mark the minima of the Stoner continuum. (c) and (d): Illustration of the splitting of the lower electron bands of the $Z_2$ Hubbard model. Insets show the positions of the corresponding local maxima (red solid circles) of the spin-up bands and local minima (purple solid circles) of the spin-down bands in the first Brillouin zone. Green arrows denote the corresponding scattering channels of the minima of the Stoner continuum marked in (a) and (b). $t_2=0.65$ for (a) and (c) and $t_2=0.741$ for (b) and (d). Other parameters are fixed at $t_1=1.0$, $U=2.3$.}
\label{tispectrum}
\end{figure}

%\par The spin-1 excitation spectrum of the the ferromagnetic phase of the topological Hubbard model is shown in Fig.~\ref{fig4}. This result indicates that the ground state of the topological Hubbard model is indeed the fully spin-polarized ferromagnetic state due to the fact that the flipping a spin from such state requires a positive energy for $U_{c}>1.82t_{1}$.

%\par Because of the spin-dependent hopping term, only $z$ component of the total spin operator commutes with the Hamiltonian, the symmetry of the Hamiltonian is explicitly broken to U(1) from SU(2) spin rotation symmetry. A Goldstone mode will not exist in the spin-wave excitation spectrum since the ground state has the same symmetry with the Hamiltonian. Since the symmetry is not spontaneously broken, generally, the spin-wave excitations should be gapped [see Fig.~\ref{fig4}]. The low energy excitations consist of two branches of collective modes identified as the spin-wave excitations degenerating along the $\Gamma-X$ path in the Brillouin Zone below the Stoner continuum. There is a critical $U_{c}=1.82t_{1}$ ($W/U_{c}=0.455$) shown in Fig.~\ref{fig4}(c), when $U<U_{c}$, the ferromagnetic ground state vanishes. Interestingly, the Stoner continuums of the Chern and topological Hubbard models are almost the same because their $G_{1}(\mathbf{k}_{i},\mathbf{Q})$ dominating the Stoner excitation gap are equal [see Fig.~\ref{fig3}(b) and Fig.~\ref{fig4}(a)]. As is shown in the Chern hubbard model, the dispersion of the $c$ band also makes the excitation spectrum have the downwarping behavior (while the collective mode (the red dashed line) is partly absorbed in the Stoner continuum) and gives rise to the phase transition for $U=1.82t_{1}$ [see Fig.~\ref{fig4}]. There exist explicit differences from the result of the bosonization formalism in the spin-wave excitation spectrum and it is not clear whether those differences come from the neglected spin-wave interaction in the bosonization formalism.

\section{Summary and discussion}\label{sd}
%\par In summary, we study the ferromagnetic phase in the 1/4-filled nearly flat-band topological Hubbard model on the square lattice in two cases. The spin-1 excitation spectrums of the models are determined by the PED method. In one case, the model is called Chern Hubbard model, whose hopping term is the $\pi$-flux model with nonzero Chern number and nearly flat-band. The model has SU(2) spin rotation symmetry and  breaks time-reversal symmetry. For $ U>U_{c}=2t_{1}$, the ground state is a completely spin-polarized ferromagnetic ground state. The low energy excitations are identified as the acoustic and optical spin-wave excitation, respectively, the high energy excitations, as the Stoner continuum. The spin-wave excitation spectrum is gapless due to spontaneous broken symmetry associated with the Goldstone mode at the center of the Brillouin Zone. In another case, the model is called topological Hubbard model, whose hopping term is the nearly flat-band $\pi$-flux model with nonzero spin Chern number. The model is invariant under time-reversal symmetry and breaks SU(2) to U(1) spin rotation symmetry. For $ U>U_{c}=1.82t_{1}$, the ground state exhibits a completely spin-polarized ferromagnetic ground state and the excitation spectrum is gapped. The low energy excitations consist of two types of collective modes identified as spin-wave excitations below the Stoner continuum.

%\par We have discussed the effects of the nearly flat-band on the excitation spectrum and the phase transition. The nearly band makes the excitation spectrum have the downwarping behavior. If we assume the nearly flat-band as a completely flat-band, the result indicates that the ground state is a fully spin-polarized ferromagnetic state no matter how small the repulsion $U$ is and there is no phase transition. The reason of phase transition is due to the fact that the nearly flat-band is not a completely flat-band. Finally, for the  Chern Hubbard model, our spin-wave excitation spectrum is qualitatively similar to the result of the bosonization formalism in ref.~\cite{Doretto} if the nearly flat-band is assumed as a completely flat-band. However, for topological Hubbard model, there are explicit differences from the result of the bosonization formalism in the spin-wave excitation spectrum and it is not clear whether those differences come from the neglected spin-wave interaction in the bosonization formalism.

\begin{acknowledgments}
\par This work was supported by the National Natural Science Foundation of China (11774152) and National Key Projects for Research and Development of China (Grant No. 2016YFA0300401).
\par X.-F. S. and Z.-L. G. contributed equally to this work.
\end{acknowledgments}

\bibliography{ref}
\end{document}

### 2018-06-08 manager.py

@@ -1,0 +1,69 @@
+import mkl
+import numpy as np
+from HamiltonianPy import *
+from source import *
+from collections import OrderedDict
+from fractions import Fraction
+
+def tbatasks(name,parameters,lattice,terms,jobs=()):
+    import HamiltonianPy.FreeSystem as TBA
+    tba=tbaconstruct(name,parameters,lattice,terms)
+    if 'EB' in jobs:
+        if len(lattice.vectors)==2:
+            tba.register(EB(name='EB',path=square_gxm(reciprocals=lattice.reciprocals,nk=100),run=TBA.TBAEB))
+        elif len(lattice.vectors)==1:
+            tba.register(EB(name='EB',path=KSpace(reciprocals=lattice.reciprocals,segments=[(-0.5,0.5)],end=True,nk=401),run=TBA.TBAEB))
+        else:
+            tba.register(EB(name='EB',run=TBA.TBAEB))
+        tba.summary()
+
+def fbfmtasks(name,parameters,lattice,terms,interactions,nk=50,scalefree=1.0,scaleint=1.0,jobs=()):
+    import HamiltonianPy.FBFM as FB
+    assert  len(lattice.vectors)==2
+    ns,ne=len(lattice),len(lattice)/2
+    basis=FB.FBFMBasis(BZ=FBZ(lattice.reciprocals,nks=(nk,nk)),filling=Fraction(ne,ns*2))
+    fbfm=fbfmconstruct(name,parameters,basis,lattice,terms,interactions)
+    if 'EB' in jobs:
+        fbfm.register(FB.EB(name='EB%s'%nk,path='S:G-Y,Y-M,M-G',ne=nk**2,scalefree=scalefree,scaleint=scaleint,plot=True,run=FB.FBFMEB))
+        fbfm.summary()
+    if 'KPOS' in jobs:
+        path=[fbfm.basis.BZ[pos] for pos in fbfm.basis.BZ.path(KMap(fbfm.lattice.reciprocals,'S:G-X,X-M,M-G'),mode='I')]
+        pos=len(path)/6*5
+        mode='D'
[…]

### 2018-06-08 paper/figure.py

@@ -1,0 +1,560 @@
+import HamiltonianPy as HP
+import numpy as np
+import itertools as it
+import matplotlib.pyplot as plt
+import matplotlib.colors as colors
+import scipy.interpolate as si
+import sys
+import pdb
+
+def lattice():
+    plt.ion()
+    fig=plt.figure()
+    gs=plt.GridSpec(1,2)
+    fig.subplots_adjust(left=0.02,right=0.98,top=0.99,bottom=0.0,hspace=0.0,wspace=0.08)
+
+    ax=fig.add_subplot(gs[0,0])
+    ax.axis('off')
+    ax.axis('equal')
+    ax.set_xlim(-0.1,2.1)
+    ax.set_ylim(-0.1,2.1)
+    P,t1,t2=np.array([0.0,0.0]),np.array([1.0,0.0]),np.array([0.0,1.0])
+    lattice=HP.Lattice(name='2DFB',rcoords=HP.tiling(cluster=[P],vectors=[t1,t2],translations=it.product(xrange(3),xrange(3))),neighbours=2)
+    for bond in lattice.bonds:
+        p1,p2=bond.spoint,bond.epoint
+        if bond.neighbour==0:
+            ax.scatter(p1.rcoord[0],p1.rcoord[1],s=100,edgecolors='none',color='blue' if p1.pid.site%2==0 else 'red',zorder=3)
+        elif bond.neighbour==1:
+            ax.plot([p1.rcoord[0],p2.rcoord[0]],[p1.rcoord[1],p2.rcoord[1]],lw=2,color='black',zorder=2)
+            theta,s1,s2=HP.azimuthd(bond.rcoord),p1.pid.site%2,p2.pid.site%2
+            if np.abs(theta)<HP.RZERO or np.abs(theta-180)<HP.RZERO:
+                coeff=1 if (s1,s2)==(0,1) else -1
+            else:
+                coeff=-1 if (s1,s2)==(0,1) else 1
+            center,direction=(p1.rcoord+p2.rcoord)/2,bond.rcoord/10*coeff
+            ax.annotate(s='',xy=center-direction/2,xytext=center+direction/2,arrowprops={'color':'black','linewidth':2,'arrowstyle':'->','zorder':3})
+        else:
[…]

### 2018-06-08 result/fbfm/compress.py

@@ -1,0 +1,9 @@
+from os import listdir
+from os.path import isfile
+import numpy as np
+
+for fname in listdir('.'):
+    if isfile(fname) and fname[-4:]=='.dat':
+        print 'Compression of %s'%fname
+        data=np.loadtxt(fname)
+        np.savez_compressed(fname[:-4],data=data)

### 2018-06-08 source/__init__.py

@@ -1,0 +1,3 @@
+from config import *
+from tba import *
+from fbfm import *

### 2018-06-08 source/config.py

@@ -1,0 +1,127 @@
+from HamiltonianPy import *
+import numpy as np
+
+__all__=['name1','name2','nneighbour','parametermap','idfmap','t1c','t1t','t2','U','S2xxy','KPOS','FBFMKPOS']
+
+# The configs of the model
+name1='2DCI'
+name2='2DTI'
+nneighbour=2
+
+# parametermap
+parametermap=None
+
+# idfmap
+idfmap=lambda pid: Fock(atom=0,norbital=1,nspin=2,nnambu=1)
+
+# nearest neighbour pi-flux hopping indexpacks for Chern insulator
+def cipi(bond):
+    theta=azimuthd(bond.rcoord)
+    s1,s2,phase=bond.spoint.pid.site%2,bond.epoint.pid.site%2,np.exp(1.0j*np.pi/4)
+    if np.abs(theta)<RZERO or np.abs(theta-180)<RZERO:
+        coeff=phase if (s1,s2)==(0,1) else (np.conjugate(phase) if (s1,s2)==(1,0) else 0.0)
+    elif np.abs(theta-90)<RZERO or np.abs(theta-270)<RZERO:
+        coeff=np.conjugate(phase) if (s1,s2)==(0,1) else (phase if (s1,s2)==(1,0) else 0.0)
+    else:
+        coeff=0.0
+    return sigma0('sp')*coeff
+
+# nearest neighbour pi-flux hopping indexpacks for topological insulator
+def tipi(bond):
+    theta=azimuthd(bond.rcoord)
+    s1,s2,phase=bond.spoint.pid.site%2,bond.epoint.pid.site%2,np.exp(1.0j*np.pi/4)
+    if np.abs(theta)<RZERO or np.abs(theta-180)<RZERO:
+        coeff=phase if (s1,s2)==(0,1) else (np.conjugate(phase) if (s1,s2)==(1,0) else 0.0)
+    elif np.abs(theta-90)<RZERO or np.abs(theta-270)<RZERO:
+        coeff=np.conjugate(phase) if (s1,s2)==(0,1) else (phase if (s1,s2)==(1,0) else 0.0)
+    else:
+        coeff=0.0
+    return FockPack(coeff,spins=(1,1))+FockPack(np.conjugate(coeff),spins=(0,0))
+
+# next to nearset neighbour hopping amplitude
+def nnnamplitude(bond):
+    theta=azimuthd(bond.rcoord)
[…]

### 2018-06-08 source/fbfm.py

@@ -1,0 +1,23 @@
+import numpy as np
+import HamiltonianPy.FBFM as FB
+from HamiltonianPy import *
+from config import *
+
+__all__=['fbfmconstruct']
+
+def fbfmconstruct(name,parameters,basis,lattice,terms,interactions,**karg):
+    fbfm=FB.FBFM(
+        dlog=           'log',
+        din=            'data',
+        dout=           'result/fbfm',
+        name=           '%s_%s_%s'%(name,lattice.name,basis.polarization),
+        parameters=     parameters,
+        map=            parametermap,
+        basis=          basis,
+        lattice=        lattice,
+        config=         IDFConfig(priority=FB.FBFM_PRIORITY,pids=lattice.pids,map=idfmap),
+        terms=          [term(**parameters) for term in terms],
+        interactions=   [term(**parameters) for term in interactions],
+        dtype=          np.complex128
+    )
+    return fbfm

### 2018-06-08 source/tba.py

@@ -1,0 +1,22 @@
+import numpy as np
+import HamiltonianPy.FreeSystem as TBA
+from HamiltonianPy import *
+from config import *
+
+__all__=['tbaconstruct']
+
+def tbaconstruct(name,parameters,lattice,terms,**karg):
+    config=IDFConfig(priority=DEFAULT_FERMIONIC_PRIORITY,pids=lattice.pids,map=idfmap)
+    tba=TBA.TBA(
+        dlog=       'log',
+        din=        'data',
+        dout=       'result/tba',
+        name=       '%s_%s'%(name,lattice.name),
+        parameters= parameters,
+        map=        parametermap,
+        lattice=    lattice,
+        config=     config,
+        terms=      [term(**parameters) for term in terms],
+        dtype=      np.complex128
+        )
+    return tba

### 2018-06-13 paper/paper.tex
Version 1.0.

@@
 \author{Jian-Xin Li}
+\email[]{jxli@nju.edu.cn}
 \affiliation{National Laboratory of Solid State Microstructures and Department of Physics, Nanjing University, Nanjing 210093, China}
@@
 For the Chern Hubbard model, the spin wave is gapless while for the $Z_2$ Hubbard model, the spin wave is gapped.
-Remarkably, for both cases, the nonflatness of the free electron bands introduces dips of the lower boundary of the Stoner continuum, \textcolor{red}{and the scatterings between the Stoner continuum at these dips and the collective modes significantly reduce the latter's energy}, which can lead to roton-like spin wave excitations.
+Remarkably, for both cases, the nonflatness of the free electron bands introduces dips of the lower boundary of the Stoner continuum, and significantly renormalizes the energies of the collective modes around these dips downward, which can lead to roton-like spin wave excitations.
 With the increase of this nonflatness, the energy of the induced roton-like modes goes down and finally touches zero, which results in the destabilization of the ferromagnetic phase.
@@
 Afterwards this ferromagnetism was shown to be stable against small nonflatness of the electron bands if and only if the Hubbard interaction exceeds a critical value \cite{T_PRL1994}.
-Spin wave excitations over this ferromagnetic ground state was also studied \cite{KA_PRL1994,SGDL_arXiv2008} and itinerant topological magnons has been reported quite recently \cite{SGDL_arXiv2008}.
+Spin wave excitations over this ferromagnetic ground state was also studied \cite{KA_PRL1994,SGDL_PRB2018} and itinerant topological magnons has been reported quite recently \cite{SGDL_PRB2018}.
 
@@
 A critical Hubbard interaction strength is found for both the Chern Hubbard model and the $Z_2$ Hubbard model, upon which the ferromagnetic phase is stable.
-Furthermore the spin-1 excitation spectra are shown to consist of collective modes (spin waves) and individual modes (Stoner continuum).
+Furthermore, the spin-1 excitation spectra are shown to consist of collective modes (spin waves) and individual modes (Stoner continuum).
 For the Chern Hubbard model, the spin wave is gapless while for the $Z_2$ Hubbard model, the spin wave is gapped.
-Remarkably, for both cases, the nonflatness of the free electron bands introduces dips of the lower boundary of the Stoner continuum, \textcolor{red}{and the scatterings between the Stoner continuum at these dips and the collective modes significantly reduce the latter's energy}, which can lead to roton-like spin wave excitations.
+Remarkably, for both cases, the nonflatness of the free electron bands introduces dips of the lower boundary of the Stoner continuum, and significantly renormalizes the energies of the collective modes around these dips downward, which can lead to roton-like spin wave excitations.
 With the increase of this nonflatness, the energy of the induced roton-like modes goes down and finally touches zero, which results in the destabilization of the ferromagnetic phase.
@@
 \subsection{Exact diagonalization with projection}\label{exact_diag_proj}
-\par Exact diagonalization method with a projection onto the low-energy Hilbert space bas been widely applied to systems that host flat or nearly-flat energy bands \cite{NSCM_PRL2011,RB_PRX2011,NSRCM_PRB2011,SGDL_arXiv2008,NSRCM_PRL2012,DG_PRB2015}.
+\par Exact diagonalization method with a projection onto the low-energy Hilbert space bas been widely applied to systems that host flat or nearly-flat energy bands \cite{NSCM_PRL2011,RB_PRX2011,NSRCM_PRB2011,SGDL_PRB2018,NSRCM_PRL2012,DG_PRB2015}.
 This approach ignores the band mixing of electrons and applies when the Coulomb interaction is small compared to the energy gap of the electron bands.
@@
 Therefore this basis labels a spin-1 scattering channel with the index $\mathbf{k}_{i}$.
-Thus the dimension of this Hilbert space scales linearly with respect to the number of electron momentums \cite{SGDL_arXiv2008}, and a much larger system can be numerically accessed than the usual exact diagonalization without projection, which enables us to analyze the properties of the spin-1 excitation spectra at the whole first Brillouin zone rather than some restricted discrete points solely.
+Thus the dimension of this Hilbert space scales linearly with respect to the number of electron momentums \cite{SGDL_PRB2018}, and a much larger system can be numerically accessed than the usual exact diagonalization without projection, which enables us to analyze the properties of the spin-1 excitation spectra at the whole first Brillouin zone rather than some restricted discrete points solely.
 The matrix element of the projected Hamiltonian on this spin-1 excitation basis can be easily obtained after some algebra,
@@
 To approach this limit, long-range hopping terms in the real space must be included \cite{NSCM_PRL2011}.
-To elaborate the physics related to the nonflatness of the $d$ band, we also calculated the spin-1 excitation spectra in the flatband limit for comparison.
+To reveal the physics related to the nonflatness of the $d$ band, we also calculated the spin-1 excitation spectra in the flatband limit for comparison.
 This can be done by simply ignoring the $\left[\varepsilon_d(\mathbf{k}_i-\mathbf{q})-\varepsilon_d(\mathbf{k}_i)\right]\delta_{\mathbf{k}_j,\mathbf{k}_i}$ term in Eq.
@@
 
-\par The phase diagrams of the quarter-filled Chern Hubbard model and $Z_2$ Hubbard model are shown in Fig.
+\par Before the detailed discussion on the spin-1 excitation spectra of the quarter-filled Chern Hubbard model and $Z_2$ Hubbard model, we present their phase diagrams first, which are shown in Fig.
 \ref{phase}.
@@
 As is discussed in Sec.
-\ref{exact_diag_proj}, the phase boundary is determined as the onset point of $U$ below which negative spin-1 excitation spectra appear.
-It is clear that a critical Hubbard interaction strength is needed to maintain the ferromagnetically ordered ground state when the electron band has finite nonflatness.
-This observation is expected as a result from the competition between the kinetic energy and the potential energy of the electrons, as a fully spin-polarized state minimizes the energy of Hubbard interactions but cost more energy when the electron band disperses.
-An intriguing fact about the phase diagrams is that the phase boundaries always lie near the contour line with $U/W=2.0$.
+\ref{exact_diag_proj}, the phase boundary is determined as the onset point of $U$ below which negative spin-1 excitation spectra arise.
+The phase diagrams of these two models appear quite similar.
+It is obvious that a critical Hubbard interaction strength is needed to maintain the ferromagnetically ordered ground state when the electron band has finite nonflatness.
+Furthermore the phase boundaries always lie near the contour line with $U/W=2.0$.
+Qualitatively speaking, these observations are expected as a result from the competition between the kinetic energy and the potential energy of the electrons, as a fully spin-polarized state minimizes the energy of Hubbard interactions but cost more energy when the electron band disperses.
+In the following, we will elaborate the spin-1 excitation spectra of these models to investigate the physics observed above microscopically and quantitatively.
 
-\subsection{Chern Hubbard model}
+\subsection{Chern Hubbard model}\label{CHM}
 \begin{figure}
@@
 \caption{(Color online) (a) Spin-1 excitation spectra of the Chern Hubbard model in the FM phase.
-Green solid lines denote the spin waves and the grey regions denote the Stoner continuum.
+Green solid lines denote the spin waves and the grey region denotes the Stoner continuum.
 Green dashed lines represent the corresponding spectra in the flatband limit.
@@
 
-\par To gain a comprehensive understanding of the physics observed above, the spin-1 excitation spectra along a high symmetry path in the first Brillouin zone of the quarter-filled Chern Hubbard model with two different $t_2$ values are plotted in Fig.
-\ref{cispectrum}(a)-(b) as the green solid lines and the shaded areas.
-For a comparison, the spectra for those in the corresponding flat-band limit are shown as the green dashed lines as well.
-Apparently, for both the original model and the flat-band limit model, the spectra consist of two types of modes: the low-lying ones exhibiting well-defined band structures and the high-energy ones forming a continuum.
-The former are identified as the magnons with an acoustic branch and an optical branch while the latter are identified as the Stoner continuum.
-At $\Gamma$ point, the acoustic magnon band is gapless, which is the character of a ferromagnetic excitation with the Goldstone mode as the result of the spontaneous spin $SU(2)$ symmetry breaking.
-Remarkably,
+\par In this subsection, we focus on the quarter-filled Chern Hubbard model.
+Its spin-1 excitation spectra in the FM phase along a high symmetry path in the first Brillouin zone are plotted in Fig.
+\ref{fmcispectrum}(a) as two low-lying green solid lines and a high-energy shaded area, with the parameters marked by the red star in Fig.
+\ref{phase}(a).
+The modes in the green solid lines and the shaded area take on distinct behaviors: the former exhibits well-defined band structures while the latter forms a continuum.
+In fact, for these two kinds of modes, the patterns of the contributions from each spin-1 particle-hole scattering channel of electrons, which are embodied in the eigenvectors $\Psi_{\mathbf{q}}(\mathbf{k}_i)$ of Eq.
+\ref{PHP}, are quite different.
+In Fig.
+\ref{fmcispectrum}(b), the spectral weights of each scattering channel, i.e.
+$|\Psi_{\mathbf{q}}(\mathbf{k}_i)|^2$ as a function of $\mathbf{k}_i$, for the four lowest eigen levels with the center-of-mass momentum $\mathbf{q}=(0,0)$ ($\Gamma$ point) are shown.
+It is clear that for the modes in the green solid lines [Fig.
+\ref{fmcispectrum}($\text{b}_1$) and Fig.
+\ref{fmcispectrum}($\text{b}_2$)], the spectral weights come from a quite broad range of scattering channels; indeed, the spectral weight for the lowest eigen mode at $\Gamma$ point is even homogeneous for all scattering channels.
+However, for those in the shaded area [Fig.
+\ref{fmcispectrum}($\text{b}_3$) and Fig.
+\ref{fmcispectrum}($\text{b}_4$)], the spectral weights almost come from a single scattering channel.
+Therefore, the former modes are collective spin-1 excitations and are identified as the spin waves while the latter modes are individual spin-1 excitations and are identified as the Stoner continuum.
+Due to the $AB$ sublattice of the model, the spin waves contains an acoustic branch and an optical branch.
+The acoustic branch is gapless, which is consistent with the Goldstone theorem as the spin-fully polarized ground state spontaneously breaks the spin $SU(2)$ rotation symmetry.
+
+\par To explore the physics related to the nonflatness of the lower electron band, the corresponding spin-1 excitation spectra in the flatband limit are also shown in Fig.
+\ref{fmcispectrum}(a) as the green dashed lines.
+The flatband-limit spectra consist of two parts as well, i.e.
+the low-lying spin waves and the high-energy Stoner continuum.
+However, in this case, the Stoner continuum is just a pile of featureless flat lines at $U/2$.
+Comparing the spectra of the original model with this, as indicated by the black arrows in Fig.
+\ref{fmcispectrum}(a), we can conclude that the nonflatness of the electron band introduces dips of the lower boundary of the Stoner continuum.
+Furthermore, around the valleys of these dips, the energies of the collective modes are apparently renormalized downward, leading to the occurrence of roton-like local minima of the optical branch.
+
+\par The above physics can be illustrated by a simple band-splitting picture as shown in Fig.
+\ref{fmcispectrum}(c).
+Roughly speaking, in the presence of ferromagnetism, the spin-down electron band and spin-up electron band are split by an energy gap proportional to $U\langle m \rangle$, where $U$ is the Hubbard interaction strength and $\langle m \rangle$ is the average magnetic moment per site.
+For a quarter-filled electron model with two inequivalent sublattices, $\langle m \rangle=1/2$.
+Thus the energy splitting is $U/2$.
+Then an individual spin-1 excitation, or a mode in the Stoner continuum, can be viewed as the excitation of an electron from the lower fully-filled spin-up band to the upper empty spin-down band.
+This argument can be verified by our numerical results.
+In fact, the sharp locality on the $\mathbf{k}_i$ space of the spectral weights of a mode in the Stoner continuum [as has be seen in Fig.
+\ref{fmcispectrum}($\text{b}_3$) and Fig.
+\ref{fmcispectrum}($\text{b}_4)$] implies that this kind of mode is mainly determined by the corresponding diagonal term in Eq.
+\ref{PHP}.
+By numerics, we also find that the term $\frac{U}{N} \sum_{a=1,2} \sum_{\mathbf{p}\neq\mathbf{k}_i} \left| \mu_{a,\mathbf{p}\uparrow} \right|^2 \left| \mu_{a,\mathbf{k}_{i}-\mathbf{q}\downarrow} \right|^2 \sim \frac{U}{2}$.
+Therefore, the excitation energies of the Stoner continuum can be approximated by $\varepsilon_{d\downarrow} (\mathbf{k}_i-\mathbf{q}) + \frac{U}{2} -\varepsilon_{d\uparrow} (\mathbf{k}_i)$, which is consistent with the above picture.
+In Fig.
+\ref{fmcispectrum}(a), the lower and upper boundaries of the Stoner continuum approximated by this argument are plotted as the blue solid lines, which fits extremely well with the true boundaries of the shaded area, thus verifies the validity of this simple argument.
+Now we can find out the dominant scattering channels that contribute most to the dips of the Stoner continuum and read off their center-of-mass momentums easily.
+In the inset of Fig.
+\ref{fmcispectrum}(c), we mark the local maxima of the spin-up band with red solid circles and the local minima of the spin-down band with purple ones.
+The scattering channels corresponding to the valley bottoms marked by the black arrows in Fig.
+\ref{fmcispectrum}(a) are also indicated by green arrows.
+Obviously, a spin-1 excitation of an electron from one of the spin-up maxima to one of the spin-down minima appears as a valley bottom of the lower boundary of the Stoner continuum and the momentum difference of these two electron state provides the center-of-mass momentum of that spin-1 excitation.
+
+\par Within this picture, the collective modes of the spectra can be viewed as bound states in the gap of the Stoner continuum.
+Then the physics of the downward renormalization of the energies of the collective modes induced by the dips of the Stoner continuum is quite similar to that observed in the two-dimensional antiferromagents, where the suppression of the energies of the spin waves are explained by their couplings to the nearby deconfined spinon continuum \cite{ZFSMC_PRL2006,TS_PRL2013,DMCNTPEMIR_NP2015,YWDYL_arXiv2018}.
+To understand the renormalization of the collective modes observed here, in Fig.
+\ref{fmcispectrum}(d), we plot the spectral weights of the four lowest eigen levels subtracted by those in the flatband limit with the center-of-mass momentum $q=(\pi,0)$ ($X_1$ point).
+The differences concentrate around the scattering channels labeled by $\mathbf{k}_i=(\pm\pi,0)$ and $\mathbf{k}_i=(0,\pm\pi)$.
+However, the results for the collective modes [Fig.
+\ref{fmcispectrum}($\text{d}_1$) and Fig.
+\ref{fmcispectrum}($\text{d}_2$)] are positive while those for the individual modes at the valley bottom of the Stoner continuum [Fig.
+\ref{fmcispectrum}($\text{d}_3$) and Fig.
+\ref{fmcispectrum}($\text{d}_4$)] are negative.
+Therefore a noticeable amount of spectral weights transfer from the latter to the former, which indicates a strong coupling between the spin waves and the nearby Stoner continuum.
+This coupling renormalizes the energies of the spin waves and leads to the occurrence of a roton-like local minimum for the optical branch.
 
@@
 
-%\par The Hamiltonian has SU(2) spin rotation symmetry.
-The ferromagnetic ground state breaks SU(2) spin rotation symmetry to U(1) symmetry, and one would therefore expect the presence of a Goldstone mode in the spin-wave excitation spectrum due to spontaneous broken symmetry.
-We show such a mode at the center of the Brillouin Zone [see Fig.~\ref{fig3}].
-The lowest two bands (black solid line and red dashed line) respectively corresponds to the acoustic and optical magnon (spin-wave excitation) bands degenerating along the $X-M$ path in the Brillouin Zone while the higher excitations represent to Stoner continuum in Fig.~\ref{fig3}.
-$G_{1}(\mathbf{k}_{i},\mathbf{Q})$ dominates the Stoner excitation gap while $G_{2}(\mathbf{k}_{i},\mathbf{k}_{j},\mathbf{Q})$ has very little contribution to the Stoner gap.
-With the increase of $W/U$, the optical spin-wave excitation is eventually absorbed in the Stoner continuum along the $\Gamma-X$ path in the Brillouin Zone [see Fig.
-\ref{fig3}(a),(b),(c)].
-For $U_{c}=2t_{1}$ ($W/U_{c}=0.414$), there is a critical situation of ferromagnetic ground state, consistent with the theory of Tasaki \cite{Tasaki1} [see Fig.~\ref{fig3}(c)].
-When $W/U>W/U_{c}$, the spin-wave excitation appears to be negative somewhere, and the ground state is no longer fully spin-polarized ferromagnetic state.
-In order to investigate the effect of the $c$ band to the spin-1 excitation spectrum, we have to consider the case that the $c$ band is assumed as a completely flat-band.
-In this case, the excitation energies are proportional to the on-site repulsion $U$ instead of the nearest-neighbor hopping energy $t_{1}$ and the boundary of Stoner continuum coalesces into a line [see Fig.~\ref{fig3}(d)].
-No matter how small the repulsion $U$ is, the ground state is a fully spin-polarized ferromagnetic state.
-In contrast to this case, the dispersion of the $c$ band makes the excitation spectrum have the downwarping behavior and gives rise to the phase transition for $U=2t_{1}$ [see Fig.~\ref{fig3}(a),(b),(c)].
-The spin-wave excitation spectrum in Fig.~\ref{fig3}(d) is qualitatively similar to the result of the bosonization formalism in ref.~\cite{Doretto}.
-However, Our result takes into account the effect of the spin-wave interaction in the bosonization formalism.
-The bosonization formalism neglects the contribution of the nearly flat-band to the spin wave excitation and can't determine what the $W/U$ value is, the ground state is the fully spin-polarized ferromagnetic state.
+\par The physics discussed above is essential for the destabilization of the ferromagnetic phase.
+In Fig.
+\ref{pbcispectrum}(a), we plot the spin-1 excitation spectra of the model at the FM/NFM phase boundary, with the parameters marked by the blue star in Fig.
+\ref{phase}(a).
+The corresponding illustration of the split-band picture is shown in Fig.
+\ref{pbcispectrum}(b).
+It can be seen that with this increased nonflatness, the valley bottom marked by the black arrow in Fig.
+\ref{pbcispectrum}(a) of the Stoner continuum goes deeper toward zero energy.
+Besides the roton-like local minimum on the optical band, a roton-like local minimum on the acoustic band also appears, which can be seen more clearly in the inset of Fig.
+\ref{pbcispectrum}(a).
+With the further increase of the nonflatness of the electron band, the energy of this newly formed roton mode will go down through the zero energy, which leads to the final destabilization of the ground state.
+From Fig.
+\ref{fmcispectrum}(c) and Fig.
+\ref{pbcispectrum}(b), we can see that the ferromagnetic ground state is quite robust against the spin-1 flips until the top of the spin-up band is approaching near to the bottom of the spin-down band.
+Considering that the energy splitting of these two bands is $U/2$, the destabilization point would be quite close to $U/2\sim W$.
 
-\subsection{Topological Hubbard model}
+\subsection{$Z_2$ Hubbard model}\label{ZHM}
 \begin{figure}
@@
 
-%\par The spin-1 excitation spectrum of the the ferromagnetic phase of the topological Hubbard model is shown in Fig.~\ref{fig4}.
-This result indicates that the ground state of the topological Hubbard model is indeed the fully spin-polarized ferromagnetic state due to the fact that the flipping a spin from such state requires a positive energy for $U_{c}>1.82t_{1}$.
-
-%\par Because of the spin-dependent hopping term, only $z$ component of the total spin operator commutes with the Hamiltonian, the symmetry of the Hamiltonian is explicitly broken to U(1) from SU(2) spin rotation symmetry.
-A Goldstone mode will not exist in the spin-wave excitation spectrum since the ground state has the same symmetry with the Hamiltonian.
-Since the symmetry is not spontaneously broken, generally, the spin-wave excitations should be gapped [see Fig.~\ref{fig4}].
-The low energy excitations consist of two branches of collective modes identified as the spin-wave excitations degenerating along the $\Gamma-X$ path in the Brillouin Zone below the Stoner continuum.
-There is a critical $U_{c}=1.82t_{1}$ ($W/U_{c}=0.455$) shown in Fig.~\ref{fig4}(c), when $U<U_{c}$, the ferromagnetic ground state vanishes.
-Interestingly, the Stoner continuums of the Chern and topological Hubbard models are almost the same because their $G_{1}(\mathbf{k}_{i},\mathbf{Q})$ dominating the Stoner excitation gap are equal [see Fig.~\ref{fig3}(b) and Fig.~\ref{fig4}(a)].
-As is shown in the Chern hubbard model, the dispersion of the $c$ band also makes the excitation spectrum have the downwarping behavior (while the collective mode (the red dashed line) is partly absorbed in the Stoner continuum) and gives rise to the phase transition for $U=1.82t_{1}$ [see Fig.~\ref{fig4}].
-There exist explicit differences from the result of the bosonization formalism in the spin-wave excitation spectrum and it is not clear whether those differences come from the neglected spin-wave interaction in the bosonization formalism.
+\par In this subsection, we focus on the quarter-filled $Z_2$ Hubbard model.
+Its spin-1 excitation spectra along a high symmetry path in the first Brillouin zone with the parameters marked by the red star and blue star in Fig.
+\ref{phase}(b) are plotted in Fig.
+\ref{tispectrum}(a) and Fig.
+\ref{tispectrum}(b), respectively.
+As is similar to Sec.
+\ref{CHM}, the low-lying green solid lines represent the spin waves (collective modes) and the high-energy shaded areas represent the Stoner continuum (individual modes).
+A distinction from the Chern Hubbard model is that the collective modes here are gapped because the $Z_2$ Hubbard model explicitly breaks the spin $SU(2)$ rotation symmetry and no spontaneous continuous symmetry breaking occurs in the ferromagnetic phase.
+The corresponding spectra in the flatband limit are shown as the green dashed lines in Fig.
+\ref{tispectrum}(a) and Fig.
+\ref{tispectrum}(b) as well for comparison.
+Similarly, the nonflatness of the electron band introduces dips of the Stoner continuum, and the energies of the spin waves around the valley bottoms of the Stoner continuum are renormalized downward, which results in roton-like modes on both branches of the spin waves.
+The dominant scattering channels leading to the dips of the Stoner continuum and the renormalization of the spin waves can also be determined by the previously proposed band-splitting picture, which is illustrated in Fig.
+\ref{tispectrum}(c) and Fig.
+\ref{tispectrum}(d).
+Their insets shows the corresponding local maxima (red solid circles) of the spin-up bands and local minima (purple solid circles) of the spin-down bands in the first Brillouin zone.
+Here only the scattering channels that contribute to the valley bottoms marked by the black arrows in Fig.
+\ref{tispectrum}(a) and Fig.
+\ref{tispectrum}(b) are stressed because their induced roton-like modes lead to the destabilization of the ferromagnetic ground state [See Fig.
+\ref{tispectrum}(b)].
+We want to remark that although the center-of-mass momentums of the forementioned valley bottoms are the same, the scattering channels are different because the bottom of the spin-down band changes from the $\Gamma$ and $M_{1,2,3,4}$ points to the $X_{1,2}$ and $Y_{1,2}$ points when $t_2$ changes from $0.65$ to $0.741$ [See Fig.
+\ref{tispectrum}(c) and Fig.
+\ref{tispectrum}(d)].
+Last but not least, the ferromagnetic ground state is also quite robust against the spin flips until the bottom of the spin-down band almost overlaps with the spin-up band on energy, which means the phase boundary of the FM/NFM is also quite close to $U/W=2$.
 
 \section{Summary and discussion}\label{sd}
-%\par In summary, we study the ferromagnetic phase in the 1/4-filled nearly flat-band topological Hubbard model on the square lattice in two cases.
-The spin-1 excitation spectrums of the models are determined by the PED method.
-In one case, the model is called Chern Hubbard model, whose hopping term is the $\pi$-flux model with nonzero Chern number and nearly flat-band.
-The model has SU(2) spin rotation symmetry and  breaks time-reversal symmetry.
-For $ U>U_{c}=2t_{1}$, the ground state is a completely spin-polarized ferromagnetic ground state.
-The low energy excitations are identified as the acoustic and optical spin-wave excitation, respectively, the high energy excitations, as the Stoner continuum.
-The spin-wave excitation spectrum is gapless due to spontaneous broken symmetry associated with the Goldstone mode at the center of the Brillouin Zone.
-In another case, the model is called topological Hubbard model, whose hopping term is the nearly flat-band $\pi$-flux model with nonzero spin Chern number.
-The model is invariant under time-reversal symmetry and breaks SU(2) to U(1) spin rotation symmetry.
-For $ U>U_{c}=1.82t_{1}$, the ground state exhibits a completely spin-polarized ferromagnetic ground state and the excitation spectrum is gapped.
-The low energy excitations consist of two types of collective modes identified as spin-wave excitations below the Stoner continuum.
+\par In summary, we have considered the flatband ferromagnetic phases in a quarter-filled Chern Hubbard model and a quarter-filled $Z_2$ Hubbard model in this paper.
+By using the numerical exact diagonalization method with a projection onto the lower nearly-flat electron band, we determine the critical Hubbard interaction strength upon which the ferromagnetic phase is stable, and elaborate the ferromagnetic spin-1 excitation spectra of these models.
+Both spectra consist of collective modes (spin waves) and individual modes (Stoner continuum).
+For the Chern Hubbard model, the spin wave is gapless while for the $Z_2$ Hubbard model, the spin wave is gapped.
+Remarkably, for both cases, the nonflatness of the free electron bands introduces dips of the lower boundary of the Stoner continuum, and significantly renormalizes the energies of the collective modes around these dips downward, which can lead to roton-like spin wave excitations.
+With the increase of this nonflatness, the energy of the induced roton-like modes goes down and finally touches zero, which results in the destabilization of the ferromagnetic phase.
 
-%\par We have discussed the effects of the nearly flat-band on the excitation spectrum and the phase transition.
-The nearly band makes the excitation spectrum have the downwarping behavior.
-If we assume the nearly flat-band as a completely flat-band, the result indicates that the ground state is a fully spin-polarized ferromagnetic state no matter how small the repulsion $U$ is and there is no phase transition.
-The reason of phase transition is due to the fact that the nearly flat-band is not a completely flat-band.
-Finally, for the  Chern Hubbard model, our spin-wave excitation spectrum is qualitatively similar to the result of the bosonization formalism in ref.~\cite{Doretto} if the nearly flat-band is assumed as a completely flat-band.
-However, for topological Hubbard model, there are explicit differences from the result of the bosonization formalism in the spin-wave excitation spectrum and it is not clear whether those differences come from the neglected spin-wave interaction in the bosonization formalism.
+\par We want to make a comparison of our flatband-limit results with those obtained by the generalized bosonization formalism proposed in Ref.
+\cite{DG_PRB2015}, where both the nonflatness of the electron bands and the self-interactions between the magnons are ignored.
+For the Chern Hubbard model, our results are qualitatively consistent with theirs [See Fig.
+\ref{fmcispectrum}(a) in our paper and Fig.
+4 in Ref.
+\cite{DG_PRB2015}].
+However, for the $Z_2$ Hubbard model, our results are different from theirs in that the band bottom of the spin waves in our results lies at the $M_{1,2,3,4}$ points while theirs lies at the $\Gamma$ point [See Figs.
+\ref{tispectrum}(a)-(b) in our paper and Fig.
+5 in Ref.
+\cite{DG_PRB2015}].
+We attribute this discrepancy to the self-interactions of the magnons because our numerical method treat the system after the projection exactly.
+Although correlated nearly-flat topological bands are hard to realize in real materials, they are possible to be designed in cold-atom systems.
+In fact great progress has been made very recently \cite{LCJPS_N2009,AALBPB_PRL2013,MSKBK_PRL2013,JMDLUGE_N2014}.
+We thus expect our results to be tested in future cold-atom experiments.
 

### 2018-06-14 paper/paper.tex
Version 1.1

@@
 It was proposed in a pioneering work by Haldane \cite{H_PRL1988} that a spinless fermionic model on a honeycomb lattice exhibits integer quantum Hall effect \cite{KDP_PRL1980} without an external magnetic field.
-This model, serving as the first example of Chern insulator, breaks the time reversal symmetry with a complex next nearest neighbor hopping and is characterized by a nonzero Chern number \cite{TKNN_PRL1982}.
+This model, serving as the first example of Chern insulator, breaks the time reversal symmetry with a complex next-nearest-neighbor hopping and is characterized by a nonzero Chern number \cite{TKNN_PRL1982}.
 Later, the concept was generalized to time reversal symmetric systems with spin-orbital coupling (SOC), such as the monolayer graphene \cite{KM_PRL2005a,KM_PRL2005b} and HgTe/CdTe quantum wells \cite{BHZ_S2006,Ketc_S2007}.
@@
 \par The lattice models with nontrivial band topology share much similarity with the two-dimensional electron gas (2DEG) under a strong magnetic field with Landau levels, e.g.
-the existence of topologically protected gapless edge states \cite{H_PRL1988,L_PRB1981,H_PRB1982,KM_PRL2005a,KM_PRL2005b,BHZ_S2006,Ketc_S2007}.
+the existence of topologically protected gapless edge states \cite{H_PRL1988,L_PRB1981,H_PRB1982,KM_PRL2005a,KM_PRL2005b,BHZ_S2006,Ketc_S2007,YXL_PRL2011,LYGL_PRB2016,LGLW_NJP2017}.
 Thus more novel phases other than the Chern insulator or $Z_2$ insulator are expected when Coulomb interactions are taken into account, as is similar to the fractional quantum Hall effect \cite{TSG_PRL1982,L_PRL1983} in the 2DEG with Landau levels.
@@
 \begin{equation}
-t_1^{ij,\sigma}=t_1\exp\left[i(-1)^{\delta^1_{ij}}\alpha_\sigma\pi/4\right],
+t_1^{ij,\sigma}=t_1\exp\left[i\delta^1_{ij}\alpha_\sigma\pi/4\right],
 \end{equation}
 \begin{equation}
-t_2^{ij}=t_2(-1)^{\delta^2_{ij}}.
+t_2^{ij}=t_2\delta^2_{ij}.
 \end{equation}
@@
 It can be seen that there exists a gap between the $d$ band and $f$ band when $t_1\ne0$ and $t_2\ne0$.
-\par When there is a gap between the $d$ band and the $f$ band, these bands can be shown to be topologically non-trivial by calculating their Chern numbers \cite{TKNN_PRL1982} (for the time-reversal-symmetry breaking case) or $Z_2$ indices \cite{KM_PRL2005b,SWSH_PRL2006} (for the time-reversal-symmetry preserving case).
+\par When there is a gap between the $d$ band and the $f$ band, these bands can be shown to be topologically non-trivial by calculating their Chern numbers \cite{TKNN_PRL1982} (for the time-reversal-symmetry-breaking case) or $Z_2$ indices \cite{KM_PRL2005b,SWSH_PRL2006} (for the time-reversal-symmetry-preserving case).
 The Chern number for a single spin component of the $d$ band or the $f$ band can be expressed in terms of the coefficients $D_{i,\mathbf{k}}$ \cite{HK_RMP2010,QZ_RMP2011},
@@
 \includegraphics[width=\columnwidth]{phase}
-\caption{(Color online) Phase diagram of the quarter-filled (a) Chern Hubbard model and (b) $Z_2$ Hubbard model.
+\caption{(Color online) Phase diagrams of the quarter-filled (a) Chern Hubbard model and (b) $Z_2$ Hubbard model.
 The colormap represents the ratio $U/W$ of the lower electron band, where $U$ is the Hubbard interaction strength and $W$ is the lower electron bandwidth.
@@
 Green dashed lines represent the corresponding spectra in the flatband limit.
-Blue solid lines denote the upper and lower boundaries of the Stoner continuum determined by the split band picture shown in (c).
+Blue solid lines denote the upper and lower boundaries of the Stoner continuum determined by the band-splitting picture shown in (c).
 Black arrows mark the local minima of the Stoner continuum.
@@
 Inset shows the positions of the corresponding local maxima (red solid circles) of the spin-up bands and local minima (purple solid circles) of the spin-down bands in the first Brillouin zone.
-Green arrows denote the corresponding scattering channels of the dips of the Stoner continuum marked in (a).
+Green arrows denote the corresponding scattering channels of the valley bottoms of the Stoner continuum marked in (a).
 (d) Differences of the spectral weights subtracted by those in the flatband limit for the lowest four eigen levels of the spin-1 excitation spectra with the center-of-mass momentum $\mathbf{q}=(\pi,0)$.
@@
 \ref{fmcispectrum}(a), we can conclude that the nonflatness of the electron band introduces dips of the lower boundary of the Stoner continuum.
-Furthermore, around the valleys of these dips, the energies of the collective modes are apparently renormalized downward, leading to the occurrence of roton-like local minima of the optical branch.
+Furthermore, around these dips, the energies of the collective modes are apparently renormalized downward, leading to the occurrence of roton-like local minima on the optical branch.
 
@@
 \ref{fmcispectrum}(c).
-Roughly speaking, in the presence of ferromagnetism, the spin-down electron band and spin-up electron band are split by an energy gap proportional to $U\langle m \rangle$, where $U$ is the Hubbard interaction strength and $\langle m \rangle$ is the average magnetic moment per site.
+Roughly speaking, in the presence of ferromagnetism, the spin-down electron band and spin-up electron band are split in energy proportional to $U\langle m \rangle$, where $U$ is the Hubbard interaction strength and $\langle m \rangle$ is the average magnetic moment per site.
 For a quarter-filled electron model with two inequivalent sublattices, $\langle m \rangle=1/2$.
@@
 \ref{fmcispectrum}(a), the lower and upper boundaries of the Stoner continuum approximated by this argument are plotted as the blue solid lines, which fits extremely well with the true boundaries of the shaded area, thus verifies the validity of this simple argument.
-Now we can find out the dominant scattering channels that contribute most to the dips of the Stoner continuum and read off their center-of-mass momentums easily.
+Now we can find out the dominant scattering channels that contribute most to the valley bottoms of the Stoner continuum and read off their center-of-mass momentums easily.
 In the inset of Fig.
@@
 \par Within this picture, the collective modes of the spectra can be viewed as bound states in the gap of the Stoner continuum.
-Then the physics of the downward renormalization of the energies of the collective modes induced by the dips of the Stoner continuum is quite similar to that observed in the two-dimensional antiferromagents, where the suppression of the energies of the spin waves are explained by their couplings to the nearby deconfined spinon continuum \cite{ZFSMC_PRL2006,TS_PRL2013,DMCNTPEMIR_NP2015,YWDYL_arXiv2018}.
+Then the physics of the downward renormalization of the energies of the collective modes induced by the valley bottoms of the Stoner continuum is quite similar to that observed in the two-dimensional antiferromagents, where the suppression of the energies of the spin waves are explained by their couplings to the nearby deconfined spinon continuum \cite{ZFSMC_PRL2006,TS_PRL2013,DMCNTPEMIR_NP2015,YWDYL_arXiv2018}.
 To understand the renormalization of the collective modes observed here, in Fig.
@@
 \ref{tispectrum}(d)].
-Last but not least, the ferromagnetic ground state is also quite robust against the spin flips until the bottom of the spin-down band almost overlaps with the spin-up band on energy, which means the phase boundary of the FM/NFM is also quite close to $U/W=2$.
+Last but not least, the ferromagnetic ground state is also quite robust against the spin flips until the bottom of the spin-down band almost overlaps with the top of the spin-up band on energy, which means the phase boundary of the FM/NFM is also quite close to $U/W=2$.
 

### 2018-06-26 paper/paper.tex
Version 1.2.

@@
 \begin{document}
-\title{Flatband ferromagnetism and dispersion relations of spin-1 excitations in topological Hubbard models}
+\title{Ferromagnetism and spin excitations in topological Hubbard models with a flat-band}
 \author{Xiao-Fei Su}
@@
 The model Hamiltonian either explicitly breaks the time-reversal symmetry but preserves the spin $SU(2)$ rotation symmetry (Chern Hubbard model), or preserves the time-reversal symmetry but explicitly breaks the spin $SU(2)$ rotation symmetry ($Z_2$ Hubbard model).
-By using the numerical exact diagonalization method with a projection onto the lower nearly-flat electron band, we determine the critical Hubbard interaction strength upon which the ferromagnetic phase is stable, and elaborate the ferromagnetic spin-1 excitation spectra of the Chern Hubbard model and the $Z_2$ Hubbard model.
+By using the numerical exact diagonalization method with a projection onto the lower nearly-flat electron band, we determine the critical Hubbard interaction upon which the ferromagnetic phase is stable, and obtain the ferromagnetic spin-1 excitation spectra of the Chern Hubbard model and the $Z_2$ Hubbard model.
 Both spectra consist of collective modes (spin waves) and individual modes (Stoner continuum).
 For the Chern Hubbard model, the spin wave is gapless while for the $Z_2$ Hubbard model, the spin wave is gapped.
-Remarkably, for both cases, the nonflatness of the free electron bands introduces dips of the lower boundary of the Stoner continuum, and significantly renormalizes the energies of the collective modes around these dips downward, which can lead to roton-like spin wave excitations.
-With the increase of this nonflatness, the energy of the induced roton-like modes goes down and finally touches zero, which results in the destabilization of the ferromagnetic phase.
+Remarkably, in both cases, the nonflatness of the free electron bands introduces dips of the lower boundary of the Stoner continuum, and significantly renormalizes the energies of the collective modes around these dips downward, which can lead to roton-like spin wave excitations.
+We elaborate that the destabilization of the ferromagnetic phase happens when the energy of the induced roton-like modes becomes zero with the increase of the nonflatness.
 \end{abstract}
@@
 Then the low energy physics is dominated by the one-spin-flip excitations.
-These spin-1 excitations has been studied by a generalized bosonization scheme where the interacting fermionic model is mapped to a free bosonic model describing spin-wave excitations at the harmonic approximation \cite{DG_PRB2015}.
+These spin-1 excitations have been studied by a generalized bosonization scheme where the interacting fermionic model is mapped to a free bosonic model describing spin-wave excitations at the harmonic approximation \cite{DG_PRB2015}.
 The ferromagnetism was shown to be stable against such spin wave excitations, which are gapless in the Chern Hubbard model and gapped in the $Z_2$ Hubbard model.
 However in this bosonization scheme, the free part of the electron model plays no role in the spin wave excitations other than contributing a global constant, which means it should fail due to the competition between the kinetic energy and potential energy of electrons when the nonflatness of the electron bands is not negligible \cite{T_PRL1994}.
-What's worse, in a strictly local periodic tight-binding model, an energy band with a nonzero Chern number cannot be exactly flat \cite{CMST_JPMT2014}.
-Therefore it remains an open question on whether the ground state is stable against the spin-1 excitations and how the the nonflatness of the electron bands manifests itself in the spin-1 excitation spectra in such models.
+In fact, in a strictly local periodic tight-binding model, an energy band with a nonzero Chern number cannot be exactly flat \cite{CMST_JPMT2014}.
+Therefore it remains an open question on whether the ground state is stable against the spin-1 excitations and how the nonflatness of the electron bands manifests itself in the spin-1 excitation spectra in such models.
 
 \par To elucidate these questions, we adopt the numerical exact diagonalization method with a projection onto the lower nearly-flat band to take close investigations on the spin-1 excitations of the models.
-A critical Hubbard interaction strength is found for both the Chern Hubbard model and the $Z_2$ Hubbard model, upon which the ferromagnetic phase is stable.
+A critical magnitude of the Hubbard interaction is found for both the Chern Hubbard model and the $Z_2$ Hubbard model, upon which the ferromagnetic phase is stable.
 Furthermore, the spin-1 excitation spectra are shown to consist of collective modes (spin waves) and individual modes (Stoner continuum).
 For the Chern Hubbard model, the spin wave is gapless while for the $Z_2$ Hubbard model, the spin wave is gapped.
-Remarkably, for both cases, the nonflatness of the free electron bands introduces dips of the lower boundary of the Stoner continuum, and significantly renormalizes the energies of the collective modes around these dips downward, which can lead to roton-like spin wave excitations.
+Remarkably, for both cases, the nonflatness of the free electron bands introduces dips of the lower boundary of the Stoner continuum, and significantly renormalizes the energies of the collective modes around these dips downward, which leads to roton-like spin wave excitations.
 With the increase of this nonflatness, the energy of the induced roton-like modes goes down and finally touches zero, which results in the destabilization of the ferromagnetic phase.
+We also make a comparison of our results in the flatband limit with those obtained by the bosonization scheme \cite{DG_PRB2015}.
+Qualitative agreements are observed for the Chern Hubbard model, yet notable discrepancy appears for the $Z_2$ Hubbard model.
+We attribute the discrepancy to the self-interactions of the magnons, which is ignored in the bosonization scheme but treated exactly in our method after the projection.
 
@@
 \cite{DG_PRB2015}].
-However, for the $Z_2$ Hubbard model, our results are different from theirs in that the band bottom of the spin waves in our results lies at the $M_{1,2,3,4}$ points while theirs lies at the $\Gamma$ point [See Figs.
+However, for the $Z_2$ Hubbard model, our results are different from theirs at two points at least [See Figs.
 \ref{tispectrum}(a)-(b) in our paper and Fig.
@@
 \cite{DG_PRB2015}].
+On the one hand, the band bottom of the spin waves in our results lies at the $M_{1,2,3,4}$ points while theirs lies at the $\Gamma$ point .
+On the other hand, the acoustic branch and optical branch are degenerate along the $\Gamma$-$X$ path in our results while they are degenerate along the $X$-$M$ path in theirs.
 We attribute this discrepancy to the self-interactions of the magnons because our numerical method treat the system after the projection exactly.

### 2018-09-18 paper/paper.tex
Version 1.3

@@
 \begin{document}
-\title{Ferromagnetism and spin excitations in topological Hubbard models with a flat-band}
+\title{Ferromagnetism and spin excitations in topological Hubbard models with a flatband}
 \author{Xiao-Fei Su}
@@
 The model Hamiltonian either explicitly breaks the time-reversal symmetry but preserves the spin $SU(2)$ rotation symmetry (Chern Hubbard model), or preserves the time-reversal symmetry but explicitly breaks the spin $SU(2)$ rotation symmetry ($Z_2$ Hubbard model).
-By using the numerical exact diagonalization method with a projection onto the lower nearly-flat electron band, we determine the critical Hubbard interaction upon which the ferromagnetic phase is stable, and obtain the ferromagnetic spin-1 excitation spectra of the Chern Hubbard model and the $Z_2$ Hubbard model.
+By using the numerical exact diagonalization method with a projection onto the lower nearly-flat electron band, we determine the critical Hubbard interaction below which the ferromagnetic phase is unstable, and obtain the ferromagnetic spin-1 excitation spectra of the Chern Hubbard model and the $Z_2$ Hubbard model.
 Both spectra consist of collective modes (spin waves) and individual modes (Stoner continuum).
@@
 This model, serving as the first example of Chern insulator, breaks the time reversal symmetry with a complex next-nearest-neighbor hopping and is characterized by a nonzero Chern number \cite{TKNN_PRL1982}.
-Later, the concept was generalized to time reversal symmetric systems with spin-orbital coupling (SOC), such as the monolayer graphene \cite{KM_PRL2005a,KM_PRL2005b} and HgTe/CdTe quantum wells \cite{BHZ_S2006,Ketc_S2007}.
+Later, the concept was generalized to time reversal symmetric systems with spin-orbital coupling (SOC), such as the graphene \cite{KM_PRL2005a,KM_PRL2005b} and HgTe/CdTe quantum wells \cite{BHZ_S2006,Ketc_S2007}.
 The SOC there generates complex hopping terms similar to that proposed by Haldane but with opposite chiralities for electrons with up spins and down spins, resulting in the quantum spin Hall insulator characterized by a $Z_2$ index.
@@
 \par Another involved intriguing phenomenon arising from Coulomb repulsions in flat or nearly-flat bands is the itinerant ferromagnetism \cite{T_PTP1998}.
-It was proved by Tasaki and Mielke that the ground state of an flat electron band with its filling factor not more than but sufficiently close to $1/2$ is ferromagnetically ordered as long as an infinitesimal onsite Hubbard repulsion is present \cite{T_PRL1992,M_PLA1993,MT_CMP1993}.
-Afterwards this ferromagnetism was shown to be stable against small nonflatness of the electron bands if and only if the Hubbard interaction exceeds a critical value \cite{T_PRL1994}.
-Spin wave excitations over this ferromagnetic ground state was also studied \cite{KA_PRL1994,SGDL_PRB2018} and itinerant topological magnons has been reported quite recently \cite{SGDL_PRB2018}.
+It was proved by Tasaki and Mielke that the ground state of a flat electron band with its filling factor not more than but sufficiently close to $1/2$ is ferromagnetically ordered as long as an infinitesimal onsite Hubbard repulsion is present \cite{T_PRL1992,M_PLA1993,MT_CMP1993}.
+Afterwards, this ferromagnetism was shown to be stable against small nonflatness of the electron bands if and only if the Hubbard interaction exceeds a critical value \cite{T_PRL1994}.
+Spin wave excitations over this ferromagnetic ground state were also studied \cite{KA_PRL1994,SGDL_PRB2018} and itinerant topological magnons have been reported quite recently \cite{SGDL_PRB2018}.
 
@@
 Furthermore, ferromagnetism can also lead to possible high-temperature quantum anomalous Hall effect (QAHE) when the nearly-flat topological band is half-filled \cite{NSRCM_PRL2012}.
-In this paper, we focus on this ferromagnetism induced QAHE in nearly-flat topological bands.
-As a paradigm, we consider a square lattice Hubbard model whose free part is the $\pi$ flux model with topologically nontrivial and nearly-flat electron bands.
+In this paper, we study the ferromagnetism and  spin excitations from the ferromagnetic ground state in nearly-flat topological bands.
+As a paradigm, we consider a square lattice Hubbard model whose free part is a $\pi$ flux model with topologically nontrivial and nearly-flat electron bands.
 The model Hamiltonian either explicitly breaks the time-reversal symmetry but preserves the spin $SU(2)$ rotation symmetry (Chern Hubbard model), or preserves the time-reversal symmetry but explicitly breaks the spin $SU(2)$ rotation symmetry ($Z_2$ Hubbard model).
 When the model is quarter filled (or correspondingly, the lower nearly-flat band is half filled), the ground state is spin fully polarized due to the ferromagnetism and exhibits QAHE because of the nonzero Chern number of a single-spin band.
-Then the low energy physics is dominated by the one-spin-flip excitations.
+Consequently, the charge excitations are gapped and the low energy physics is dominated by the one-spin-flip excitations.
 These spin-1 excitations have been studied by a generalized bosonization scheme where the interacting fermionic model is mapped to a free bosonic model describing spin-wave excitations at the harmonic approximation \cite{DG_PRB2015}.
 The ferromagnetism was shown to be stable against such spin wave excitations, which are gapless in the Chern Hubbard model and gapped in the $Z_2$ Hubbard model.
-However in this bosonization scheme, the free part of the electron model plays no role in the spin wave excitations other than contributing a global constant, which means it should fail due to the competition between the kinetic energy and potential energy of electrons when the nonflatness of the electron bands is not negligible \cite{T_PRL1994}.
+However in this bosonization scheme, the free part of the electron model plays no role in the spin wave excitations other than contributing a global constant, suggesting that it should fail due to the competition between the kinetic energy and potential energy of electrons when the nonflatness of the electron bands is not negligible \cite{T_PRL1994}.
 In fact, in a strictly local periodic tight-binding model, an energy band with a nonzero Chern number cannot be exactly flat \cite{CMST_JPMT2014}.
@@
 \par To elucidate these questions, we adopt the numerical exact diagonalization method with a projection onto the lower nearly-flat band to take close investigations on the spin-1 excitations of the models.
-A critical magnitude of the Hubbard interaction is found for both the Chern Hubbard model and the $Z_2$ Hubbard model, upon which the ferromagnetic phase is stable.
+A critical magnitude of the Hubbard interaction is found for both the Chern Hubbard model and the $Z_2$ Hubbard model, below which the ferromagnetic phase is unstable.
 Furthermore, the spin-1 excitation spectra are shown to consist of collective modes (spin waves) and individual modes (Stoner continuum).
@@
 Qualitative agreements are observed for the Chern Hubbard model, yet notable discrepancy appears for the $Z_2$ Hubbard model.
-We attribute the discrepancy to the self-interactions of the magnons, which is ignored in the bosonization scheme but treated exactly in our method after the projection.
+We attribute the discrepancy to the self-interactions of the magnons, which are ignored in the bosonization scheme but treated exactly in our method after the projection.
 
@@
 In Sec.
-\ref{mm}, we introduce the Chern Hubbard model and $Z_2$ Hubbard model studied in this paper, discuss the nontrivial band topology of their free parts, interpret the emergence of QAHE resulting from the interplay of flat-band ferromagnetism and nontrivial band topology, and formulate the exact diagonalization method with a projection onto the lower nearly-flat band on details.
+\ref{mm}, we introduce the Chern Hubbard model and $Z_2$ Hubbard model studied in this paper, discuss the nontrivial band topology of their free parts, interpret the emergence of QAHE resulting from the interplay between flatband ferromagnetism and nontrivial band topology, and formulate the exact diagonalization method with a projection onto the lower nearly-flat band on details.
 In Sec.
@@
 Blue and red solid circles denote the $A$ and $B$ sublattices, respectively.
-The nearest-neighbor hopping amplitudes (solid black lines) are equal to $t_{1}\exp[i\alpha(\sigma)\pi/4]$ ($\alpha(\sigma)=1$ for the time-reversal-symmetry-breaking case and $\alpha(\sigma)=\pm1$ for the time-reversal-symmetry-preserving case) along the direction of the arrows, the next-nearest-neighbor hopping amplitudes are equal to $t_{2}$ (dashed green lines) and $-t_{2}$ (dashed purple lines).
+The nearest-neighbor hopping amplitudes (solid black lines) are equal to $t_{1}\exp(i\alpha_{\sigma}\pi/4)$ (see text) along the direction of the arrows, the next-nearest-neighbor hopping amplitudes are equal to $t_{2}$ (dashed green lines) and $-t_{2}$ (dashed purple lines).
 The shaded area denotes the unit cell.
 (b) First Brillouin Zone.
-$\Gamma=(0,0)$, $X_{1,2}=(\pm\pi,0)$, $Y_{1,2}=(0,\pm\pi)$, $M_{1,2,3,4}=(\pm\pi,\pm\pi)$.}
+$\Gamma=(0,0)$, $X_{1,2}=(\pm\pi,0)$, $Y_{1,2}=(0,\pm\pi)$, $M_{1,2,3,4}=(\pm\pi,\pm\pi)$, $O_{1,2,3,4}=(\pm\frac{\pi}{2},\pm\frac{\pi}{2})$.}
 \label{lattice}
@@
 
-\par We consider a generalized $\pi$-flux Hubbard model on the square lattice, whose Hamiltonian can be written as $H=H_0+H_U$, where $H_0$ is the generalization with electron spin of the original spinless model proposed in Ref.
-\cite{NSCM_PRL2011},
+\par We consider a generalized $\pi$-flux Hubbard model on the square lattice, whose Hamiltonian can be written as $H=H_0+H_U$, where $H_0$ is the spinfull genralization of the original spinless model proposed in Ref.
+\onlinecite{NSCM_PRL2011},
 \begin{equation}
@@
 \end{equation}
-Here, $c_{i\sigma}^{\dagger}(c_{i\sigma})$ creates (annihilates) a spin $\sigma$ electron at site $i$, $n_{i\sigma}$ is the spin $\sigma$ electron number operator at site $i$, $\langle ij\rangle$ denotes the nearest-neighbor (NN) bonds and $\langle\langle ij\rangle\rangle$ denotes the next-nearest-neighbor (NNN) bonds.
-As is shown in Fig.
+Here, $c_{i\sigma}^{\dagger}(c_{i\sigma})$ creates (annihilates) a spin $\sigma$ electron at site $i$, $n_{i\sigma}=c^{\dag}_{i\sigma}c_{i\sigma}$ is the particle-number operator, $\langle ij\rangle$ denotes the nearest-neighbor (NN) bonds and $\langle\langle ij\rangle\rangle$ denotes the next-nearest-neighbor (NNN) bonds.
+As shown in Fig.
 \ref{lattice}(a), the spin-dependent NN hopping amplitude $t_1^{ij,\sigma}$ and the spin-independent NNN hopping amplitude $t_2^{ij}$ are given by
 \begin{equation}
-t_1^{ij,\sigma}=t_1\exp\left[i\delta^1_{ij}\alpha_\sigma\pi/4\right],
+t_1^{ij,\sigma}=t_1\exp\left(i\delta_1^{ij}\alpha_\sigma\pi/4\right)
 \end{equation}
+and
 \begin{equation}
-t_2^{ij}=t_2\delta^2_{ij}.
+t_2^{ij}=t_2\delta_2^{ij},
 \end{equation}
-Here, $\delta^1_{ij}=+1$ if the NN electron hopping is along the direction of the solid black arrow and $\delta^1_{ij}=-1$ if along the reversed direction.
-$\delta^2_{ij}=+1$ if the NNN electron hopping is along the dashed green lines and $\delta^2_{ij}=-1$ if along the dashed purple lines.
-The spin-dependent phase $\alpha_\sigma$ breaks the time-reversal symmetry but preserves the spin $SU(2)$ rotation symmetry if $\alpha_\uparrow=\alpha_\downarrow=+1$, whereas it preserves the time-reversal symmetry but breaks the spin $SU(2)$ rotation symmetry if $\alpha_\uparrow=+1$ and $\alpha_\downarrow=-1$.
+respectively.
+Here, $\delta_1^{ij}=+1$ if the NN electron hopping is along the direction of the solid black arrow and $\delta_1^{ij}=-1$ if along the reversed direction.
+$\delta_2^{ij}=+1$ if the NNN electron hopping is along the dashed green lines and $\delta_2^{ij}=-1$ if along the dashed purple lines.
+The spin-dependent phase factor $\alpha_\sigma$ breaks the time-reversal symmetry but preserves the spin $SU(2)$ rotation symmetry if $\alpha_\uparrow=\alpha_\downarrow=+1$, whereas it preserves the time-reversal symmetry but breaks the spin $SU(2)$ rotation symmetry if $\alpha_\uparrow=+1$ and $\alpha_\downarrow=-1$.
 
@@
 Therefore, $H_0$ describes free electrons hopping on a square lattice in the presence of a fictitious staggered $\pi$-flux pattern \cite{WWZ_PRB1989}.
-For the time-reversal-symmetry-breaking case, $\alpha_\uparrow=\alpha_\downarrow$, the flux experienced by spin-up electrons and spin-down electrons is the same, while for the time-reversal-symmetry-preserving case, $\alpha_\uparrow=-\alpha_\downarrow$, it is opposite.
+For the time-reversal-symmetry-breaking case, $\alpha_\uparrow=\alpha_\downarrow$, the fluxes experienced by spin-up electrons and spin-down electrons are the same, while for the time-reversal-symmetry-preserving case, $\alpha_\uparrow=-\alpha_\downarrow$, they are opposite.
 
-\subsection{Topology of free part}\label{topo_free_part}
-\par Gapped noninteracting fermionic systems can be topologically classified in the presence of symmetries by their Hamiltonians in the momentum space\cite{SRFL_PRB2008}.
+\subsection{Topology of free term $H_0$}\label{topo_free_part}
+\par Gapped noninteracting fermionic systems can be topologically classified by their Hamiltonians in the momentum space in the presence of symmetries \cite{SRFL_PRB2008}.
 After the Fourier transformation, the free part $H_0$ of our model reads
@@
 It can be seen that there exists a gap between the $d$ band and $f$ band when $t_1\ne0$ and $t_2\ne0$.
-\par When there is a gap between the $d$ band and the $f$ band, these bands can be shown to be topologically non-trivial by calculating their Chern numbers \cite{TKNN_PRL1982} (for the time-reversal-symmetry-breaking case) or $Z_2$ indices \cite{KM_PRL2005b,SWSH_PRL2006} (for the time-reversal-symmetry-preserving case).
+\par When there is a gap between the $d$ band and $f$ band, these bands can be shown to be topologically non-trivial by calculating their Chern numbers \cite{TKNN_PRL1982} (for the time-reversal-symmetry-breaking case) or $Z_2$ indices \cite{KM_PRL2005b,SWSH_PRL2006} (for the time-reversal-symmetry-preserving case).
 The Chern number for a single spin component of the $d$ band or the $f$ band can be expressed in terms of the coefficients $D_{i,\mathbf{k}}$ \cite{HK_RMP2010,QZ_RMP2011},
@@
 of the $d$ band is $1$ and nontrivial.
-As a consequence, the the ground state of $H_0$ will be a noninteracting $Z_2$ insulator and exhibits quantum spin Hall effect(QSHE) when the lower $d$ band is fully filled.
+As a consequence, the ground state of $H_0$ will be a noninteracting $Z_2$ insulator and exhibits quantum spin Hall effect(QSHE) when the lower $d$ band is fully filled.
 
@@
 At any fractional filling, the ground state of such a system will be a trivial metal.
-Intriguingly, when the Coulomb interactions between electrons step in, the physics of the nontrivial band topology becomes more involved, especially when the band is nearly flat so that the effects of the Coulomb interactions are highly enhanced.
+Intriguingly, when the Coulomb interactions between electrons is introduced, the physics of the nontrivial band topology becomes more involved, especially when the band is nearly flat so that the effects of the Coulomb interactions are highly enhanced.
 Combined with strong Coulomb interactions, nontrivial topological phases can emerge from fractionally filled topological bands.
 In this article, we are interested in half-filled strongly-correlated nearly-flat topological bands where QAHE can arise\cite{NSCM_PRL2011,DG_PRB2015}.
-The essence for the occurrence of this nontrivial phase is the emergence of itinerant ferromagnetism on nearly-flat bands\cite{T_PTP1998}, which fully polarizes all the electron spins.
-Therefore only one spin component of the topological band will be fully filled exactly, which leads to QAHE due to the nonzero Chern number of that spin component of the band.
+The essence for the occurrence of this nontrivial phase is the emergence of itinerant ferromagnetism on nearly-flat bands\cite{T_PTP1998}, which fully polarizes all electron spins.
+Therefore, only one spin component of the topological band will be fully filled exactly, which leads to QAHE due to the nonzero Chern number of that spin component of the band.
 
 \par The Chern Hubbard model and $Z_2$ Hubbard model on square lattice described above serve as the paradigm, where half filling of the lower electron band corresponds to quarter filling of the whole system because of the existence of $AB$ sublattices.
-When $t_2/t_1$ takes values in a selected region, the $d$ band and $f$ band are quite flat in that the flatness ratio $\Delta/W$, which is defined as the ratio between the electron gap $\Delta$ between these two bands and the bandwidth $W$ of the lower band, can be as large as $4.83$ \cite{NSCM_PRL2011,DG_PRB2015}.
-In Ref.
-\cite{NSCM_PRL2011}, the quarter filled $Z_2$ Hubbard model was studied and all spin excitations were shown to be gapped in the flatband limit, which is essential for a possible high-temperature realization of QAHE.
-In Ref.
-\cite{DG_PRB2015}, both models at quarter filling were considered and a generalized bosonization scheme was developed to study the spin-1 excitation spectra.
-However, in their bosonization scheme, the free part $H_0$ of the Hamiltonian only contributes a constant to the spectra, thus it cannot capture the physics caused by the nonflatness of the topological electron bands, which is unavoidable because an energy band with a nonzero Chern number in a strictly local periodic tight binding model cannot be exactly flat \cite{CMST_JPMT2014}.
-The spin-1 excitations, as the dominant low-energy excitations, deserves closer investigations to understand the stability of such phases and the physics related to the nonflatness of the topological bands.
-In the next subsection, we will introduce the exact diagonalization method with a projection onto the lower nearly-flat electron band to elucidate these questions.
+When $t_2/t_1$ takes values in a selected region, the $d$ band and $f$ band are quite flat in that the flatness ratio $\Delta/W$, which is defined as the ratio of the gap $\Delta$ between these two bands to the bandwidth $W$ of the lower band, can be as large as $4.83$ \cite{NSCM_PRL2011,DG_PRB2015}.
+In the next subsection, we will introduce the exact diagonalization method with a projection onto the lower nearly-flat electron band to study the spin-1 excitations of the ferromagnetic phases in these two topological Hubbard models.
 
 \subsection{Exact diagonalization with projection}\label{exact_diag_proj}
-\par Exact diagonalization method with a projection onto the low-energy Hilbert space bas been widely applied to systems that host flat or nearly-flat energy bands \cite{NSCM_PRL2011,RB_PRX2011,NSRCM_PRB2011,SGDL_PRB2018,NSRCM_PRL2012,DG_PRB2015}.
-This approach ignores the band mixing of electrons and applies when the Coulomb interaction is small compared to the energy gap of the electron bands.
+\par Exact diagonalization method with a projection onto the low-energy Hilbert space has been widely applied to systems that host flat or nearly-flat energy bands \cite{NSCM_PRL2011,RB_PRX2011,NSRCM_PRB2011,SGDL_PRB2018,NSRCM_PRL2012,DG_PRB2015}.
+This approach applies when the energy gap between the flat or nearly-flat band and other bands is larger than the Coulomb interaction.
 
@@
 \end{widetext}
-Let $|\text{FM}\rangle$ denote the spin-up fully polarized state on the $d$ band,
+Let $|\text{FM}\rangle$ denotes the spin-up fully polarized state on the $d$ band,
 \begin{equation}
@@
 \ref{lattice}(b)] and $|0\rangle$ is the fermion vaccum.
+Here, $d^{\dag}_{\mathbf{k}\uparrow}$ creates a spin-up electron with momentum $\mathbf{k}$.
 Then the basis of the spin-1 excitations with a center-of-mass momentum $\mathbf{q}$ over this reference state can be written as
 \begin{equation}
-|\mathbf{k}_i\rangle_\mathbf{q}=d^{\dagger}_{\mathbf{k}_{i}-\mathbf{q}\downarrow}d_{\mathbf{k}_{i}\uparrow}|\text{FM}\rangle.
+|\mathbf{k}_i\rangle_\mathbf{q}=d^{\dagger}_{\mathbf{k}_{i}-\mathbf{q}\downarrow}d_{\mathbf{k}_{i}\uparrow}|\text{FM}\rangle,
 \end{equation}
-Here $d^{\dagger}_{\mathbf{k}_{i}-\mathbf{q}\downarrow}$ creates a spin-down electron with momentum $\mathbf{k}_{i}-\mathbf{q}$ and $d_{\mathbf{k}_{i}\uparrow}$ annihilates a spin-up electron with momentum $\mathbf{k}_{i}$.
-Therefore this basis labels a spin-1 scattering channel with the index $\mathbf{k}_{i}$.
-Thus the dimension of this Hilbert space scales linearly with respect to the number of electron momentums \cite{SGDL_PRB2018}, and a much larger system can be numerically accessed than the usual exact diagonalization without projection, which enables us to analyze the properties of the spin-1 excitation spectra at the whole first Brillouin zone rather than some restricted discrete points solely.
+which labels a spin-1 scattering channel with the index $\mathbf{k}_{i}$.
+The dimension of this Hilbert space scales linearly with respect to the number of electron momentums \cite{SGDL_PRB2018}, so a much larger system can be numerically accessed than the usual exact diagonalization without projection.
+It enables us to analyze the properties of the spin-1 excitation spectra in detail in the whole first Brillouin zone rather than some restricted discrete points solely.
 The matrix element of the projected Hamiltonian on this spin-1 excitation basis can be easily obtained after some algebra,
@@
 (\ref{PHP}).
-It is noted that $|\text{FM}\rangle$ is the true ground state only if the whole spin-1 excitation spectra have no negative eigen energies.
-Thus we can use this as the criterion to determine the destabilization point of the ferromagnetic phase.
-\par We also want to give some remarks on the flatband limit in the framework of this method. The free part $h^\text{flat}_{\mathbf{k}\sigma}$ of the Hamiltonian in the flatband limit is defined as
+It is noted that $|\text{FM}\rangle$ is the true ground state only if the whole spin-1 excitation spectra have no negative energies.
+Thus we can use this as the criterion to determine the destabilization of the ferromagnetic phase.
+\par We also want to give some remarks on the flatband limit in the framework of this method.
+The flatband-limit Hamiltonian shares the same single-particle eigenfunctions as well as the interaction terms with the original one, yet it has exactly-flat single-particle energy bands\cite{NSCM_PRL2011}.
+Thus the free part $h^\text{flat}_{\mathbf{k}\sigma}$ of the flatband-limit Hamiltonian can be defined as
 \begin{equation}
@@
 \end{equation}
-$h^\text{flat}_{\mathbf{k}\sigma}$ shares the same eigenfunctions with $h_{\mathbf{k}\sigma}$ but has exactly-flat energy bands with the eigenvalues being $\pm 1$.
-$h^\text{flat}_{\mathbf{k}\sigma}$, together with the Hubbard interaction $H_U$, defines the flatband limit of the original Hamiltonian.
+Apparently, the eigenfunctions of $h^\text{flat}_{\mathbf{k}\sigma}$ are the same with $h_{\mathbf{k}\sigma}$'s but its energy bands are exactly flat with the eigenvalues being $\pm 1$.
 To approach this limit, long-range hopping terms in the real space must be included \cite{NSCM_PRL2011}.
 To reveal the physics related to the nonflatness of the $d$ band, we also calculated the spin-1 excitation spectra in the flatband limit for comparison.
-This can be done by simply ignoring the $\left[\varepsilon_d(\mathbf{k}_i-\mathbf{q})-\varepsilon_d(\mathbf{k}_i)\right]\delta_{\mathbf{k}_j,\mathbf{k}_i}$ term in Eq.
-(\ref{PHP}), because of the same eigenfunctions shared by $h^\text{flat}_{\mathbf{k}\sigma}$ and $h_{\mathbf{k}\sigma}$.
+This can be done by simply throwing away the $\left[\varepsilon_d(\mathbf{k}_i-\mathbf{q})-\varepsilon_d(\mathbf{k}_i)\right]\delta_{\mathbf{k}_j,\mathbf{k}_i}$ term in Eq.
+(\ref{PHP}), because of the same single-particle eigenfunctions shared by the flatband-limit Hamiltonian and the original one.
 
@@
 The colormap represents the ratio $U/W$ of the lower electron band, where $U$ is the Hubbard interaction strength and $W$ is the lower electron bandwidth.
-Red stars in (a) and (b) mark the parameters used in Fig.
-\ref{fmcispectrum} and Fig.
-\ref{tispectrum}(a), respectively.
-Blue stars in (a) and (b) mark the parameters used in Fig.
-\ref{pbcispectrum}.
-and Fig.
-\ref{tispectrum}(b), respectively.}
+Red star in (a) marks the parameter used in Figs.
+\ref{fmcispectrum} and \ref{bspicture}, blue stars in (a) mark the parameters used in Fig.
+\ref{pbcispectrum}, green stars in (b) mark the parameters used in Fig.
+\ref{tispectrum}.}
 \label{phase}
@@
 \ref{phase}.
-Here, the colormap represents the ratio of the Hubbard interaction strength $U$ over the lower electron bandwidth $W$.
+Here, the colormap represents the ratio of the Hubbard interaction strength $U$ to the lower electron bandwidth $W$.
 FM denotes the ferromagnetic phase and NFM denotes the non-ferromagnetic phase.
-As is discussed in Sec.
+As discussed in Sec.
 \ref{exact_diag_proj}, the phase boundary is determined as the onset point of $U$ below which negative spin-1 excitation spectra arise.
@@
 It is obvious that a critical Hubbard interaction strength is needed to maintain the ferromagnetically ordered ground state when the electron band has finite nonflatness.
-Furthermore the phase boundaries always lie near the contour line with $U/W=2.0$.
-Qualitatively speaking, these observations are expected as a result from the competition between the kinetic energy and the potential energy of the electrons, as a fully spin-polarized state minimizes the energy of Hubbard interactions but cost more energy when the electron band disperses.
-In the following, we will elaborate the spin-1 excitation spectra of these models to investigate the physics observed above microscopically and quantitatively.
+Furthermore, the phase boundaries always lie near the contour line with $U/W=2.0$.
+Qualitatively speaking, these observations are expected as a result from the competition between the kinetic energy and potential energy of the electrons, as a fully spin-polarized state minimizes the energy of Hubbard interactions but cost more energy when the electron band disperses.
+In the following, we will elaborate the spin-1 excitation spectra of these models.
 
@@
 Green dashed lines represent the corresponding spectra in the flatband limit.
-Blue solid lines denote the upper and lower boundaries of the Stoner continuum determined by the band-splitting picture shown in (c).
+Blue solid lines denote the upper and lower boundaries of the Stoner continuum determined by the band-splitting picture shown in Fig.
+\ref{bspicture}(a).
 Black arrows mark the local minima of the Stoner continuum.
-(b) Spectral weights for the lowest four eigen levels of the spin-1 excitation spectra with the center-of-mass momentum $\mathbf{q}=(0,0)$.
-(c) Illustration of the splitting of the lower electron bands for up spins (blue solid line) and down spins (blue dashed line) in the presence of ferromagnetism of the Chern Hubbard model.
-Inset shows the positions of the corresponding local maxima (red solid circles) of the spin-up bands and local minima (purple solid circles) of the spin-down bands in the first Brillouin zone.
-Green arrows denote the corresponding scattering channels of the valley bottoms of the Stoner continuum marked in (a).
-(d) Differences of the spectral weights subtracted by those in the flatband limit for the lowest four eigen levels of the spin-1 excitation spectra with the center-of-mass momentum $\mathbf{q}=(\pi,0)$.
-The parameters are $t_1=1.0$, $t_2=0.49$ and $U=2.3$.}
+(b$_1$-b$_4$) Spectral weights for the lowest four eigen levels of the spin-1 excitation spectra with $\mathbf{q}=(0,0)$.
+The parameters are $t_1=1.0$, $t_2=0.490$ and $U=2.3$.}
 \label{fmcispectrum}
@@
 \par In this subsection, we focus on the quarter-filled Chern Hubbard model.
-Its spin-1 excitation spectra in the FM phase along a high symmetry path in the first Brillouin zone are plotted in Fig.
-\ref{fmcispectrum}(a) as two low-lying green solid lines and a high-energy shaded area, with the parameters marked by the red star in Fig.
+In Fig.
+\ref{fmcispectrum}(a), we present the spin-1 excitation spectra in the FM phase along a high symmetry path in the first Brillouin zone with the parameters marked by the red star in Fig.
 \ref{phase}(a).
-The modes in the green solid lines and the shaded area take on distinct behaviors: the former exhibits well-defined band structures while the latter forms a continuum.
+We find that there are two kinds of obviously different excitations: one contains two low-lying modes labeled by the green solid lines exhibiting well defined band structures,  and the other forms a high-energy continuum as labeled by the shaded area.
 In fact, for these two kinds of modes, the patterns of the contributions from each spin-1 particle-hole scattering channel of electrons, which are embodied in the eigenvectors $\Psi_{\mathbf{q}}(\mathbf{k}_i)$ of Eq.
-\ref{PHP}, are quite different.
+(\ref{PHP}), are quite different.
 In Fig.
 \ref{fmcispectrum}(b), the spectral weights of each scattering channel, i.e.
-$|\Psi_{\mathbf{q}}(\mathbf{k}_i)|^2$ as a function of $\mathbf{k}_i$, for the four lowest eigen levels with the center-of-mass momentum $\mathbf{q}=(0,0)$ ($\Gamma$ point) are shown.
+$|\Psi_{\mathbf{q}}(\mathbf{k}_i)|^2$ as a function of $\mathbf{k}_i$, for the four lowest levels with the center-of-mass momentum $\mathbf{q}=(0,0)$ ($\Gamma$ point) are shown.
 It is clear that for the modes in the green solid lines [Fig.
 \ref{fmcispectrum}($\text{b}_1$) and Fig.
-\ref{fmcispectrum}($\text{b}_2$)], the spectral weights come from a quite broad range of scattering channels; indeed, the spectral weight for the lowest eigen mode at $\Gamma$ point is even homogeneous for all scattering channels.
+\ref{fmcispectrum}($\text{b}_2$)], the spectral weights come from a quite broad range of scattering channels; indeed, the spectral weight for the lowest mode at $\Gamma$ point is even homogeneous for all scattering channels.
 However, for those in the shaded area [Fig.
@@
 \ref{fmcispectrum}($\text{b}_4$)], the spectral weights almost come from a single scattering channel.
-Therefore, the former modes are collective spin-1 excitations and are identified as the spin waves while the latter modes are individual spin-1 excitations and are identified as the Stoner continuum.
-Due to the $AB$ sublattice of the model, the spin waves contains an acoustic branch and an optical branch.
-The acoustic branch is gapless, which is consistent with the Goldstone theorem as the spin-fully polarized ground state spontaneously breaks the spin $SU(2)$ rotation symmetry.
+Therefore, the modes in the green solid lines are collective spin-1 excitations and are identified as the spin waves, while the modes in the shaded area are individual spin-1 excitations and are identified as the Stoner continuum.
 
-\par To explore the physics related to the nonflatness of the lower electron band, the corresponding spin-1 excitation spectra in the flatband limit are also shown in Fig.
-\ref{fmcispectrum}(a) as the green dashed lines.
-The flatband-limit spectra consist of two parts as well, i.e.
-the low-lying spin waves and the high-energy Stoner continuum.
-However, in this case, the Stoner continuum is just a pile of featureless flat lines at $U/2$.
-Comparing the spectra of the original model with this, as indicated by the black arrows in Fig.
-\ref{fmcispectrum}(a), we can conclude that the nonflatness of the electron band introduces dips of the lower boundary of the Stoner continuum.
-Furthermore, around these dips, the energies of the collective modes are apparently renormalized downward, leading to the occurrence of roton-like local minima on the optical branch.
+\par For comparison, the corresponding spin-1 excitation spectra in the flatband limit are also shown in Fig.
+\ref{fmcispectrum}(a) as the green dashed lines, which consist of low-lying spin waves and high-energy Stoner continuum as well.
+For both the original case and the flatband-limit case, the spin waves consist of an acoustic branch and an optical branch due to the $AB$ sublattice of the model.
+The acoustic branch is gapless at the $\Gamma$ point and disperses quadratically away from but near the $\Gamma$ point, which is the character of a ferromagnetic excitation with a Goldstone mode as the spin-fully polarized ground state spontaneously breaks the spin $SU(2)$ rotation symmetry.
+Additionally, the acoustic branch and the optical branch are degenerate along the $X_1$-$M_1$ path for both cases.
+
+\par We also want to compare our results with that obtained by the generalized bosonization method at the harmonic approximation [see Fig.
+4 in Ref.
+\cite{DG_PRB2015}].
+Their result can be regarded as dispersions of free magnons, as in their scheme both the nonflatness of the $d$ band and the self-interactions between magnons are ignored.
+Apparently, their result is qualitatively consistent with ours in the flatband limit.
+However, considerable distinct features arise between theirs as well as ours in the flatband limit and ours in the dispersive case: excitation energies of the spin waves around the $X_1$ and $O_1$ points are strongly suppressed and new roton-like spin-wave excitations emerge.
+Furthermore, the boundaries of the Stoner continuum for the original model is dispersive and the local minima of its lower boundary coincides with the spin waves with suppressed energies, as indicated by the black arrows in Fig.
+\ref{fmcispectrum}(a).
+Note that in the flatband limit, only the nonflatness of the $d$ band is ignored, but the self-interactions between the magnons are treated exactly.
+Therefore, the new emerged roton-like spin waves must be attributed to the nonflatness of the $d$ band.
+
+\begin{figure}
+\includegraphics[width=\columnwidth]{bspicture}
+\caption{(Color online) (a) Illustration of the splitting of the lower electron bands for up spins (blue solid line) and down spins (blue dashed line) in the presence of ferromagnetism of the Chern Hubbard model.
+(b) Positions of the local maxima (red solid circles) of the spin-up bands and local minima (purple solid circles) of the spin-down bands in the first Brillouin zone.
+Green arrows denote the corresponding scattering channels of the minima of the Stoner continuum marked in Fig.
+\ref{fmcispectrum}(a).
+(c$_1$-c$_4$) Spectral weights subtracted by those in the flatband limit for the lowest four eigen levels of the spin-1 excitation spectra with $\mathbf{q}=(\pi,0)$.}
+\label{bspicture}
+\end{figure}
 
 \par The above physics can be illustrated by a simple band-splitting picture as shown in Fig.
-\ref{fmcispectrum}(c).
+\ref{bspicture}(a).
 Roughly speaking, in the presence of ferromagnetism, the spin-down electron band and spin-up electron band are split in energy proportional to $U\langle m \rangle$, where $U$ is the Hubbard interaction strength and $\langle m \rangle$ is the average magnetic moment per site.
@@
 This argument can be verified by our numerical results.
-In fact, the sharp locality on the $\mathbf{k}_i$ space of the spectral weights of a mode in the Stoner continuum [as has be seen in Fig.
+In fact, the $\delta$ function-like distribution of the spectral weights of a mode in the Stoner continuum [as has be seen in Fig.
 \ref{fmcispectrum}($\text{b}_3$) and Fig.
-\ref{fmcispectrum}($\text{b}_4)$] implies that this kind of mode is mainly determined by the corresponding diagonal term in Eq.
-\ref{PHP}.
+\ref{fmcispectrum}($\text{b}_4)$] implies that the excitation energy of such a mode can be approximated by the corresponding diagonal term in Eq.
+(\ref{PHP}).
 By numerics, we also find that the term $\frac{U}{N} \sum_{a=1,2} \sum_{\mathbf{p}\neq\mathbf{k}_i} \left| \mu_{a,\mathbf{p}\uparrow} \right|^2 \left| \mu_{a,\mathbf{k}_{i}-\mathbf{q}\downarrow} \right|^2 \sim \frac{U}{2}$.
-Therefore, the excitation energies of the Stoner continuum can be approximated by $\varepsilon_{d\downarrow} (\mathbf{k}_i-\mathbf{q}) + \frac{U}{2} -\varepsilon_{d\uparrow} (\mathbf{k}_i)$, which is consistent with the above picture.
+Therefore, the excitation energies of the Stoner continuum are roughly $\varepsilon_{d\downarrow} (\mathbf{k}_i-\mathbf{q}) + \frac{U}{2} -\varepsilon_{d\uparrow} (\mathbf{k}_i)$, which is consistent with the above picture.
 In Fig.
 \ref{fmcispectrum}(a), the lower and upper boundaries of the Stoner continuum approximated by this argument are plotted as the blue solid lines, which fits extremely well with the true boundaries of the shaded area, thus verifies the validity of this simple argument.
-Now we can find out the dominant scattering channels that contribute most to the valley bottoms of the Stoner continuum and read off their center-of-mass momentums easily.
-In the inset of Fig.
-\ref{fmcispectrum}(c), we mark the local maxima of the spin-up band with red solid circles and the local minima of the spin-down band with purple ones.
-The scattering channels corresponding to the valley bottoms marked by the black arrows in Fig.
+Now we can find out the dominant scattering channels that contribute most to the local boundary bottoms of the Stoner continuum and read off their center-of-mass momentums easily.
+In Fig.
+\ref{bspicture}(b), we mark the local maxima of the spin-up band with red solid circles and the local minima of the spin-down band with purple ones.
+The scattering channels corresponding to the local bottoms marked by the black arrows in Fig.
 \ref{fmcispectrum}(a) are also indicated by green arrows.
-Obviously, a spin-1 excitation of an electron from one of the spin-up maxima to one of the spin-down minima appears as a valley bottom of the lower boundary of the Stoner continuum and the momentum difference of these two electron state provides the center-of-mass momentum of that spin-1 excitation.
-
-\par Within this picture, the collective modes of the spectra can be viewed as bound states in the gap of the Stoner continuum.
-Then the physics of the downward renormalization of the energies of the collective modes induced by the valley bottoms of the Stoner continuum is quite similar to that observed in the two-dimensional antiferromagents, where the suppression of the energies of the spin waves are explained by their couplings to the nearby deconfined spinon continuum \cite{ZFSMC_PRL2006,TS_PRL2013,DMCNTPEMIR_NP2015,YWDYL_arXiv2018}.
+Obviously, a spin-1 excitation corresponding to the particle-hole scattering between a spin-up maximum and a spin-down minimum appears as a local boundary bottom of the Stoner continuum and the momentum difference of these two electron state provides the center-of-mass momentum of that spin-1 excitation.
 To understand the renormalization of the collective modes observed here, in Fig.
-\ref{fmcispectrum}(d), we plot the spectral weights of the four lowest eigen levels subtracted by those in the flatband limit with the center-of-mass momentum $q=(\pi,0)$ ($X_1$ point).
+\ref{bspicture}(c), we plot the spectral weights of the four lowest levels subtracted by those in the flatband limit with the center-of-mass momentum $q=(\pi,0)$ ($X_1$ point).
 The differences concentrate around the scattering channels labeled by $\mathbf{k}_i=(\pm\pi,0)$ and $\mathbf{k}_i=(0,\pm\pi)$.
 However, the results for the collective modes [Fig.
-\ref{fmcispectrum}($\text{d}_1$) and Fig.
-\ref{fmcispectrum}($\text{d}_2$)] are positive while those for the individual modes at the valley bottom of the Stoner continuum [Fig.
-\ref{fmcispectrum}($\text{d}_3$) and Fig.
-\ref{fmcispectrum}($\text{d}_4$)] are negative.
+\ref{bspicture}($\text{c}_1$) and Fig.
+\ref{bspicture}($\text{c}_2$)] are positive while those for the individual modes at the Stoner continuum boundary bottom [Fig.
+\ref{bspicture}($\text{c}_3$) and Fig.
+\ref{bspicture}($\text{c}_4$)] are negative.
 Therefore a noticeable amount of spectral weights transfer from the latter to the former, which indicates a strong coupling between the spin waves and the nearby Stoner continuum.
@@
 \includegraphics[width=\columnwidth]{pbcispectrum}
-\caption{(Color online) (a) Spin-1 excitation spectra of the Chern Hubbard model at the FM/NFM phase boundary.
+\caption{(Color online) (a$_1$) and (b$_1$): Spin-1 excitation spectra of the Chern Hubbard model near the FM/NFM phase boundary.
 Black arrow marks the minimum of the Stoner continuum.
-Inset shows its low-energy part along $\Gamma$-$X$ path with an amplified resolution.
-(b) Illustration of the splitting of the lower electron bands of the Chern Hubbard model.
-Inset shows the positions of the corresponding local maxima (red solid circles) of the spin-up bands and local minima (purple solid circles) of the spin-down bands in the first Brillouin zone.
-Green arrows denote the corresponding scattering channels of the minimum of the Stoner continuum marked in (a).
-The parameters are $t_1=1.0$, $t_2=0.43$ and $U=2.3$.}
+(a$_2$) and (b$_2$): Low-energy parts of the spin-1 excitation spectra shown in (a$_1$) and (b$_1$) along $\Gamma$-$X_1$ path with an amplified resolution.
+(a$_3$) and (b$_3$): Illustration of the splitting of the corresponding lower electron bands.
+$t_2=0.437$ for (a) and $t_2=0.430$ for (b).
+Other parameters are fixed at $t_1=1.0$, $U=2.3$.}
 \label{pbcispectrum}
@@
 
-\par The physics discussed above is essential for the destabilization of the ferromagnetic phase.
+\par The emerged roton-like spin waves are essential for the destabilization of the ferromagnetic phase.
 In Fig.
-\ref{pbcispectrum}(a), we plot the spin-1 excitation spectra of the model at the FM/NFM phase boundary, with the parameters marked by the blue star in Fig.
+\ref{pbcispectrum}($\text{a}_1$) and \ref{pbcispectrum}($\text{b}_1$), we plot the spin-1 excitation spectra of the model near the FM/NFM phase boundary, with the parameters marked by the blue stars in Fig.
 \ref{phase}(a).
-The corresponding illustration of the split-band picture is shown in Fig.
-\ref{pbcispectrum}(b).
-It can be seen that with this increased nonflatness, the valley bottom marked by the black arrow in Fig.
-\ref{pbcispectrum}(a) of the Stoner continuum goes deeper toward zero energy.
-Besides the roton-like local minimum on the optical band, a roton-like local minimum on the acoustic band also appears, which can be seen more clearly in the inset of Fig.
-\ref{pbcispectrum}(a).
-With the further increase of the nonflatness of the electron band, the energy of this newly formed roton mode will go down through the zero energy, which leads to the final destabilization of the ground state.
+The corresponding illustrations of the split-band picture are shown in Fig.
+\ref{pbcispectrum}($\text{a}_3$) and \ref{pbcispectrum}($\text{b}_3$).
+Combined with Fig.
+\ref{fmcispectrum}(a), it can be seen that increased nonflatness leads the moving of the boundary bottom marked by $q_1$ toward zero energy.
+Furthermore, roton-like local minima on the acoustic band appears nearby, which can be seen more clearly in Fig.
+\ref{pbcispectrum}($\text{a}_2$) and \ref{pbcispectrum}($\text{b}_2$).
+With the increase of the nonflatness of the electron band, the energy of this newly formed roton mode goes down and touches zero, which leads to the destabilization of the ground state.
 From Fig.
-\ref{fmcispectrum}(c) and Fig.
-\ref{pbcispectrum}(b), we can see that the ferromagnetic ground state is quite robust against the spin-1 flips until the top of the spin-up band is approaching near to the bottom of the spin-down band.
+\ref{pbcispectrum}($\text{a}_3$) and \ref{pbcispectrum}($\text{b}_3$), we can see that the ferromagnetic ground state is quite robust against the spin-1 flips until the top of the spin-up band is approaching near to the bottom of the spin-down band.
 Considering that the energy splitting of these two bands is $U/2$, the destabilization point would be quite close to $U/2\sim W$.
@@
 \includegraphics[width=\columnwidth]{tispectrum}
-\caption{(Color online) (a) and (b): Spin-1 excitation spectra of the $Z_2$ Hubbard model.
-Green solid lines denote the spin waves and the grey regions denote the Stoner continuum.
+\caption{(Color online) (a$_1$) and (b$_1$): Spin-1 excitation spectra of the $Z_2$ Hubbard model.
+Green solid lines denote the spin waves and grey regions denote the Stoner continuum.
 Green dashed lines represent the corresponding spectra in the flatband limit.
 Black arrows mark the minima of the Stoner continuum.
-(c) and (d): Illustration of the splitting of the lower electron bands of the $Z_2$ Hubbard model.
-Insets show the positions of the corresponding local maxima (red solid circles) of the spin-up bands and local minima (purple solid circles) of the spin-down bands in the first Brillouin zone.
-Green arrows denote the corresponding scattering channels of the minima of the Stoner continuum marked in (a) and (b).
-$t_2=0.65$ for (a) and (c) and $t_2=0.741$ for (b) and (d).
-Other parameters are fixed at $t_1=1.0$, $U=2.3$.}
+(a$_2$) and (b$_2$): Illustration of the splitting of the lower electron bands of the $Z_2$ Hubbard model.
+(a$_3$) and (b$_3$): Positions of the corresponding local maxima (red solid circles) of the spin-up bands and local minima (purple solid circles) of the spin-down bands in the first Brillouin zone.
+Green arrows denote the corresponding scattering channels of the minima of the Stoner continuum marked in (a$_1$) and (b$_1$).
+$t_2=0.650$ for (a) and $t_2=0.741$ for (b).
+Other parameters are fixed at $t_1=1.0$, $U=2.0$.}
 \label{tispectrum}
@@
 \par In this subsection, we focus on the quarter-filled $Z_2$ Hubbard model.
-Its spin-1 excitation spectra along a high symmetry path in the first Brillouin zone with the parameters marked by the red star and blue star in Fig.
+Its spin-1 excitation spectra along a high symmetry path in the first Brillouin zone with the parameters marked by the green stars in Fig.
 \ref{phase}(b) are plotted in Fig.
-\ref{tispectrum}(a) and Fig.
-\ref{tispectrum}(b), respectively.
+\ref{tispectrum}($\text{a}_1$) and Fig.
+\ref{tispectrum}($\text{b}_1$).
 As is similar to Sec.
 \ref{CHM}, the low-lying green solid lines represent the spin waves (collective modes) and the high-energy shaded areas represent the Stoner continuum (individual modes).
-A distinction from the Chern Hubbard model is that the collective modes here are gapped because the $Z_2$ Hubbard model explicitly breaks the spin $SU(2)$ rotation symmetry and no spontaneous continuous symmetry breaking occurs in the ferromagnetic phase.
-The corresponding spectra in the flatband limit are shown as the green dashed lines in Fig.
-\ref{tispectrum}(a) and Fig.
-\ref{tispectrum}(b) as well for comparison.
-Similarly, the nonflatness of the electron band introduces dips of the Stoner continuum, and the energies of the spin waves around the valley bottoms of the Stoner continuum are renormalized downward, which results in roton-like modes on both branches of the spin waves.
-The dominant scattering channels leading to the dips of the Stoner continuum and the renormalization of the spin waves can also be determined by the previously proposed band-splitting picture, which is illustrated in Fig.
-\ref{tispectrum}(c) and Fig.
-\ref{tispectrum}(d).
-Their insets shows the corresponding local maxima (red solid circles) of the spin-up bands and local minima (purple solid circles) of the spin-down bands in the first Brillouin zone.
-Here only the scattering channels that contribute to the valley bottoms marked by the black arrows in Fig.
-\ref{tispectrum}(a) and Fig.
-\ref{tispectrum}(b) are stressed because their induced roton-like modes lead to the destabilization of the ferromagnetic ground state [See Fig.
-\ref{tispectrum}(b)].
-We want to remark that although the center-of-mass momentums of the forementioned valley bottoms are the same, the scattering channels are different because the bottom of the spin-down band changes from the $\Gamma$ and $M_{1,2,3,4}$ points to the $X_{1,2}$ and $Y_{1,2}$ points when $t_2$ changes from $0.65$ to $0.741$ [See Fig.
-\ref{tispectrum}(c) and Fig.
-\ref{tispectrum}(d)].
+One distinction from the Chern Hubbard model is that the collective modes here are gapped because the $Z_2$ Hubbard model explicitly breaks the spin $SU(2)$ rotation symmetry and no spontaneous continuous symmetry breaking occurs in the ferromagnetic phase.
+The corresponding spectra in the flatband limit are also shown in Fig.
+\ref{tispectrum}($\text{a}_1$) and Fig.
+\ref{tispectrum}($\text{b}_1$) as the green dashed lines.
+Noticeable differences can be found when we compare our result in the flatband limit with that obtained by the generalized bosonization method [see Fig.
+5 in Ref.
+\cite{DG_PRB2015}].
+On the one hand, the band bottom of the spin waves in our results lies at the $M_{1,2,3,4}$ points while theirs lies at the $\Gamma$ point.
+On the other hand, the acoustic branch and optical branch are degenerate along the $\Gamma$-$X$ path in our results while they are degenerate along the $X$-$M$ path in theirs.
+These discrepancies can be attributed to the self-interactions of the magnons which are ignored in their scheme but are treated exactly in our method.
+The above observation implies that to get correct dispersion relations, self-interactions between magnons should not be neglected for the $Z_2$ Hubbard model.
+
+\par Similar to the Chern Hubbard model, the nonflatness of the $d$ band introduces dips of the Stoner continuum, and the energies of the spin waves around the boundary bottoms of the Stoner continuum are renormalized downward, which results in roton-like modes on both branches of the spin waves.
+The dominant scattering channels leading to the dips of the Stoner continuum and the renormalization of the spin waves can also be determined by the previously proposed band-splitting picture, which are illustrated in Fig.
+\ref{tispectrum}($\text{a}_2$) and Fig.
+\ref{tispectrum}($\text{b}_2$).
+Figure \ref{tispectrum}($\text{a}_3$) and Figure \ref{tispectrum}($\text{b}_3$) show the corresponding local maxima (red solid circles) of the spin-up bands and local minima (purple solid circles) of the spin-down bands in the first Brillouin zone.
+Here, only the scattering channels that contribute to the boundary bottoms marked by the black arrows in Fig.
+\ref{tispectrum}($\text{a}_1$) and Fig.
+\ref{tispectrum}($\text{b}_1$) are presented as their induced roton-like modes lead to the destabilization of the ferromagnetic ground state [See Fig.
+\ref{tispectrum}($\text{b}_1$)].
+We want to remark that although the center-of-mass momentums of the forementioned boundary bottoms are the same, the scattering channels are different because the bottom of the spin-down band changes from the $\Gamma$/$M_{1,2,3,4}$ points to the $X_{1,2}$/$Y_{1,2}$ points when $t_2$ changes from $0.650$ to $0.741$ [See Fig.
+\ref{tispectrum}($\text{a}_2$) and Fig.
+\ref{tispectrum}($\text{b}_2$)].
 Last but not least, the ferromagnetic ground state is also quite robust against the spin flips until the bottom of the spin-down band almost overlaps with the top of the spin-up band on energy, which means the phase boundary of the FM/NFM is also quite close to $U/W=2$.
@@
 \par In summary, we have considered the flatband ferromagnetic phases in a quarter-filled Chern Hubbard model and a quarter-filled $Z_2$ Hubbard model in this paper.
-By using the numerical exact diagonalization method with a projection onto the lower nearly-flat electron band, we determine the critical Hubbard interaction strength upon which the ferromagnetic phase is stable, and elaborate the ferromagnetic spin-1 excitation spectra of these models.
+By using the numerical exact diagonalization method with a projection onto the lower nearly-flat electron band, we determine the critical Hubbard interaction strength below which the ferromagnetic phase is unstable, and elaborate the ferromagnetic spin-1 excitation spectra of these models.
 Both spectra consist of collective modes (spin waves) and individual modes (Stoner continuum).
@@
 
-\par We want to make a comparison of our flatband-limit results with those obtained by the generalized bosonization formalism proposed in Ref.
-\cite{DG_PRB2015}, where both the nonflatness of the electron bands and the self-interactions between the magnons are ignored.
-For the Chern Hubbard model, our results are qualitatively consistent with theirs [See Fig.
-\ref{fmcispectrum}(a) in our paper and Fig.
-4 in Ref.
-\cite{DG_PRB2015}].
-However, for the $Z_2$ Hubbard model, our results are different from theirs at two points at least [See Figs.
-\ref{tispectrum}(a)-(b) in our paper and Fig.
-5 in Ref.
-\cite{DG_PRB2015}].
-On the one hand, the band bottom of the spin waves in our results lies at the $M_{1,2,3,4}$ points while theirs lies at the $\Gamma$ point .
-On the other hand, the acoustic branch and optical branch are degenerate along the $\Gamma$-$X$ path in our results while they are degenerate along the $X$-$M$ path in theirs.
-We attribute this discrepancy to the self-interactions of the magnons because our numerical method treat the system after the projection exactly.
+\par The collective modes of the spin-1 excitation spectra in the gap of the Stoner continuum can be viewed as bound states of the individual spin-1 excitations.
+Therefore the physics of the downward renormalization of the energies of the collective modes induced by the boundary bottoms of the Stoner continuum may be quite similar to that observed in the two-dimensional antiferromagents, if we interpret spin waves as bound states of confined spinons and attribute the suppression of the energies of the spin waves to their couplings to the nearby deconfined spinon continuum there\cite{ZFSMC_PRL2006,TS_PRL2013,DMCNTPEMIR_NP2015,SQCCMS_PRX2017,YWDYL_arXiv2018}.
 Although correlated nearly-flat topological bands are hard to realize in real materials, they are possible to be designed in cold-atom systems.

### 2018-09-18 paper/figure.py

@@ -390,7 +390,7 @@ def pbcispectrum():
         ax.set_xlim(0,posx)
         ax.set_ylim(0,0.04)
         ax.set_xticks([0,posx/2.0,posx])
-        ax.set_xticklabels(['$\Gamma$','$q$','$X$'])
+        ax.set_xticklabels(['$\Gamma$','$q$','$X_1$'])
         ax.set_yticks([0.0,0.02,0.04])
         ax.set_yticklabels(['0','0.02','0.04'])
         for tick in ax.get_xticklabels():

### 2018-10-30 paper/paper.tex
arXiv version.

@@
 \begin{abstract}
-\par We study the spin-1 excitation spectra of the flatband ferromagnetic phases in topological Hubbard models.
-As a paradigm, we consider a quarter filled square lattice Hubbard model whose free part is the $\pi$ flux model with topologically nontrivial and nearly-flat electron bands.
-The model Hamiltonian either explicitly breaks the time-reversal symmetry but preserves the spin $SU(2)$ rotation symmetry (Chern Hubbard model), or preserves the time-reversal symmetry but explicitly breaks the spin $SU(2)$ rotation symmetry ($Z_2$ Hubbard model).
-By using the numerical exact diagonalization method with a projection onto the lower nearly-flat electron band, we determine the critical Hubbard interaction below which the ferromagnetic phase is unstable, and obtain the ferromagnetic spin-1 excitation spectra of the Chern Hubbard model and the $Z_2$ Hubbard model.
-Both spectra consist of collective modes (spin waves) and individual modes (Stoner continuum).
-For the Chern Hubbard model, the spin wave is gapless while for the $Z_2$ Hubbard model, the spin wave is gapped.
-Remarkably, in both cases, the nonflatness of the free electron bands introduces dips of the lower boundary of the Stoner continuum, and significantly renormalizes the energies of the collective modes around these dips downward, which can lead to roton-like spin wave excitations.
-We elaborate that the destabilization of the ferromagnetic phase happens when the energy of the induced roton-like modes becomes zero with the increase of the nonflatness.
+\par We study the spin-1 excitation spectra of the flatband ferromagnetic phases in interacting topological insulators.
+As a paradigm, we consider a quarter filled square lattice Hubbard model whose free part is the $\pi$ flux state with topologically nontrivial and nearly-flat electron bands, which can realize either the Chern or $Z_2$ Hubbard model.
+By using the numerical exact diagonalization method with a projection onto the nearly-flat band, we obtain the ferromagnetic spin-1 excitation spectra for both the Chern and $Z_2$ Hubbard models, consisting of spin waves and Stoner continuum.
+The spectra exhibit quite distinct dispersions for both cases, in particular the spin wave is gapless for the Chern Hubbard model, while gapped for the $Z_2$ Hubbard model.
+Remarkably, in both cases, the nonflatness of the free electron bands introduces dips in the lower boundary of the Stoner continuum.
+It significantly renormalizes the energies of the spin waves around these dips downward and leads to roton-like spin excitations.
+We elaborate that it is the softening of the roton-like modes that destabilizes the ferromagnetic phase, and determine the parameter region where the ferromagnetic phase is stable.
 \end{abstract}
@@
 \par Another involved intriguing phenomenon arising from Coulomb repulsions in flat or nearly-flat bands is the itinerant ferromagnetism \cite{T_PTP1998}.
-It was proved by Tasaki and Mielke that the ground state of a flat electron band with its filling factor not more than but sufficiently close to $1/2$ is ferromagnetically ordered as long as an infinitesimal onsite Hubbard repulsion is present \cite{T_PRL1992,M_PLA1993,MT_CMP1993}.
+It was proved by Tasaki and Mielke that the ground state of a flat electron band with a filling factor not more than but sufficiently close to $1/2$ is ferromagnetically ordered as long as an infinitesimal onsite Hubbard repulsion is present \cite{T_PRL1992,M_PLA1993,MT_CMP1993}.
 Afterwards, this ferromagnetism was shown to be stable against small nonflatness of the electron bands if and only if the Hubbard interaction exceeds a critical value \cite{T_PRL1994}.
@@
 As a paradigm, we consider a square lattice Hubbard model whose free part is a $\pi$ flux model with topologically nontrivial and nearly-flat electron bands.
-The model Hamiltonian either explicitly breaks the time-reversal symmetry but preserves the spin $SU(2)$ rotation symmetry (Chern Hubbard model), or preserves the time-reversal symmetry but explicitly breaks the spin $SU(2)$ rotation symmetry ($Z_2$ Hubbard model).
+Depending on the nearest-neighbor hopping, the model Hamiltonian either explicitly breaks the time-reversal symmetry but preserves the spin $SU(2)$ rotation symmetry (Chern Hubbard model), or preserves the time-reversal symmetry but explicitly breaks the spin $SU(2)$ rotation symmetry ($Z_2$ Hubbard model).
 When the model is quarter filled (or correspondingly, the lower nearly-flat band is half filled), the ground state is spin fully polarized due to the ferromagnetism and exhibits QAHE because of the nonzero Chern number of a single-spin band.
@@
 \par To elucidate these questions, we adopt the numerical exact diagonalization method with a projection onto the lower nearly-flat band to take close investigations on the spin-1 excitations of the models.
-A critical magnitude of the Hubbard interaction is found for both the Chern Hubbard model and the $Z_2$ Hubbard model, below which the ferromagnetic phase is unstable.
+A critical magnitude of the Hubbard interaction is found for both the Chern Hubbard model and $Z_2$ Hubbard model, below which the ferromagnetic phase is unstable.
 Furthermore, the spin-1 excitation spectra are shown to consist of collective modes (spin waves) and individual modes (Stoner continuum).
@@
 With the increase of this nonflatness, the energy of the induced roton-like modes goes down and finally touches zero, which results in the destabilization of the ferromagnetic phase.
+Therefore, we elaborate the mechanism of the instability of this flat-band ferromagnetism as the softening of the emergent roton-like modes with the increase of noflatness.
 We also make a comparison of our results in the flatband limit with those obtained by the bosonization scheme \cite{DG_PRB2015}.
 Qualitative agreements are observed for the Chern Hubbard model, yet notable discrepancy appears for the $Z_2$ Hubbard model.
-We attribute the discrepancy to the self-interactions of the magnons, which are ignored in the bosonization scheme but treated exactly in our method after the projection.
+We attribute the discrepancy to the multimagnon processes, which are ignored in the bosonization scheme but treated exactly in our method after the projection.
 
@@
 In Sec.
-\ref{nr}, we summarize the phase diagram and elaborate the spin-1 excitation spectra of both models.
+\ref{nr}, we discuss the phase diagram and elaborate the spin-1 excitation spectra of both models.
 Section \ref{sd} provides a summary and discussion.
@@
 The shaded area denotes the unit cell.
-(b) First Brillouin Zone.
-$\Gamma=(0,0)$, $X_{1,2}=(\pm\pi,0)$, $Y_{1,2}=(0,\pm\pi)$, $M_{1,2,3,4}=(\pm\pi,\pm\pi)$, $O_{1,2,3,4}=(\pm\frac{\pi}{2},\pm\frac{\pi}{2})$.}
+(b) The first Brillouin zone.
+The next-nearest-neighbor distance is set to one, so $\Gamma=(0,0)$, $X=(\pi,0)$, $Y=(0,\pi)$, $M=(\pi,\pi)$, $O=(\frac{\pi}{2},\frac{\pi}{2})$.}
 \label{lattice}
@@
 \end{equation}
-Here, $c_{i\sigma}^{\dagger}(c_{i\sigma})$ creates (annihilates) a spin $\sigma$ electron at site $i$, $n_{i\sigma}=c^{\dag}_{i\sigma}c_{i\sigma}$ is the particle-number operator, $\langle ij\rangle$ denotes the nearest-neighbor (NN) bonds and $\langle\langle ij\rangle\rangle$ denotes the next-nearest-neighbor (NNN) bonds.
+Here, $c_{i\sigma}^{\dagger}(c_{i\sigma})$ creates (annihilates) a spin $\sigma$ electron at site $i$, $n_{i\sigma}=c^{\dag}_{i\sigma}c_{i\sigma}$ is the particle-number operator, $\langle ij\rangle$ denotes the nearest-neighbor (NN) bonds and $\langle\langle ij\rangle\rangle$ the next-nearest-neighbor (NNN) bonds.
 As shown in Fig.
@@
 Green solid lines denote the spin waves and the grey region denotes the Stoner continuum.
-Green dashed lines represent the corresponding spectra in the flatband limit.
 Blue solid lines denote the upper and lower boundaries of the Stoner continuum determined by the band-splitting picture shown in Fig.
@@
 (b$_1$-b$_4$) Spectral weights for the lowest four eigen levels of the spin-1 excitation spectra with $\mathbf{q}=(0,0)$.
-The parameters are $t_1=1.0$, $t_2=0.490$ and $U=2.3$.}
+The parameters are $t_1=1.0$, $t_2=0.490$ and $U=2.3$.
+Green dashed lines in (a) represent the corresponding spectra in the flatband limit.}
 \label{fmcispectrum}
@@
 \ref{phase}(a).
-We find that there are two kinds of obviously different excitations: one contains two low-lying modes labeled by the green solid lines exhibiting well defined band structures,  and the other forms a high-energy continuum as labeled by the shaded area.
-In fact, for these two kinds of modes, the patterns of the contributions from each spin-1 particle-hole scattering channel of electrons, which are embodied in the eigenvectors $\Psi_{\mathbf{q}}(\mathbf{k}_i)$ of Eq.
+It is shown that there are two kinds of obviously different excitations: one contains two low-lying modes labeled by the green solid lines exhibiting well defined band structures,  and the other forms a high-energy continuum as labeled by the shaded area.
+We find that, for these two kinds of modes, the patterns of the contributions from each spin-1 particle-hole scattering channel of electrons, which are embodied in the eigenvectors $\Psi_{\mathbf{q}}(\mathbf{k}_i)$ of Eq.
 (\ref{PHP}), are quite different.
@@
 Therefore, the modes in the green solid lines are collective spin-1 excitations and are identified as the spin waves, while the modes in the shaded area are individual spin-1 excitations and are identified as the Stoner continuum.
+The spin waves consist of an acoustic branch and an optical branch due to the $AB$ sublattice of the model.
+The acoustic branch is gapless at the $\Gamma$ point and disperses quadratically away from the $\Gamma$ point, which is the character of a ferromagnetic excitation with a Goldstone mode as the spin-fully polarized ground state spontaneously breaks the spin $SU(2)$ rotation symmetry.
+Additionally, the acoustic and optical branches are degenerate along the $X$-$M$ path.
 
-\par For comparison, the corresponding spin-1 excitation spectra in the flatband limit are also shown in Fig.
-\ref{fmcispectrum}(a) as the green dashed lines, which consist of low-lying spin waves and high-energy Stoner continuum as well.
-For both the original case and the flatband-limit case, the spin waves consist of an acoustic branch and an optical branch due to the $AB$ sublattice of the model.
-The acoustic branch is gapless at the $\Gamma$ point and disperses quadratically away from but near the $\Gamma$ point, which is the character of a ferromagnetic excitation with a Goldstone mode as the spin-fully polarized ground state spontaneously breaks the spin $SU(2)$ rotation symmetry.
-Additionally, the acoustic branch and the optical branch are degenerate along the $X_1$-$M_1$ path for both cases.
-
-\par We also want to compare our results with that obtained by the generalized bosonization method at the harmonic approximation [see Fig.
+\par In the flatband limit, the spin-1 excitation spectra as shown in Fig.
+\ref{fmcispectrum}(a) as the green dashed lines show several similarities with the dispersive case, including the constitution of the spin waves and Stoner continuum exhibiting as a flat line at $E=0.5U$, the gapless acoustic spin wave and its quadratical dispersion  near the $\Gamma$ point, and the degeneracy between the acoustic and optical branches along the $X$-$M$ path.
+Comparing results in the flatband limit with those obtained by the generalized bosonization method at the harmonic approximation [see Fig.
 4 in Ref.
-\cite{DG_PRB2015}].
-Their result can be regarded as dispersions of free magnons, as in their scheme both the nonflatness of the $d$ band and the self-interactions between magnons are ignored.
-Apparently, their result is qualitatively consistent with ours in the flatband limit.
-However, considerable distinct features arise between theirs as well as ours in the flatband limit and ours in the dispersive case: excitation energies of the spin waves around the $X_1$ and $O_1$ points are strongly suppressed and new roton-like spin-wave excitations emerge.
-Furthermore, the boundaries of the Stoner continuum for the original model is dispersive and the local minima of its lower boundary coincides with the spin waves with suppressed energies, as indicated by the black arrows in Fig.
+\cite{DG_PRB2015}], one will find that two results are qualitatively quite consistent.
+In the scheme adopted in Ref.
+\cite{DG_PRB2015}, the ignorance of the interactions between magnons suggests that only the single-magnon excitations are captured.
+So, this consistence indicates that the interactions between magnons shows a negligible effect in the flatband limit.
+However, when the free band becomes dispersive, considerable distinct features appears.
+Now, the spin-wave energies around the $X=(0,\pi)$ and $O=(\pi/2,\pi/2)$ points are strongly suppressed, as a result two local minima are formed at $X$ and $O$ which can be ascribed to the emergence of the roton-like spin-wave excitations.
+Furthermore, the continuum exhibiting as a straight line in the flatband limit extends to a large grey region shown in Fig.
+\ref{fmcispectrum}(a) with dispersive boundaries, and the local minima of its lower boundary coincide with those in the spin-wave dispersion, as indicated by the black arrows in Fig.
 \ref{fmcispectrum}(a).
-Note that in the flatband limit, only the nonflatness of the $d$ band is ignored, but the self-interactions between the magnons are treated exactly.
-Therefore, the new emerged roton-like spin waves must be attributed to the nonflatness of the $d$ band.
 
@@
 
-\par The above physics can be illustrated by a simple band-splitting picture as shown in Fig.
+\par This observation shows that it is the extending of the Stoner continuum arising from the nonflatness of the $d$ band that pushes down the spin-wave excitations.
+In particular, it suggests that the couplings between the individual excitations and spin waves would be responsible for the emergence of the roton-like spin waves.
+
+\par The formation of the dispersive boundary of the Stoner continuum can be understood by a simple band-splitting picture as shown in Fig.
 \ref{bspicture}(a).
-Roughly speaking, in the presence of ferromagnetism, the spin-down electron band and spin-up electron band are split in energy proportional to $U\langle m \rangle$, where $U$ is the Hubbard interaction strength and $\langle m \rangle$ is the average magnetic moment per site.
+In the presence of ferromagnetism, the spin-down and spin-up electron bands are split in energy which is roughly proportional to $U\langle m \rangle$, with $U$ the Hubbard interaction and $\langle m \rangle$ the average magnetic moment per site.
 For a quarter-filled electron model with two inequivalent sublattices, $\langle m \rangle=1/2$.
-Thus the energy splitting is $U/2$.
-Then an individual spin-1 excitation, or a mode in the Stoner continuum, can be viewed as the excitation of an electron from the lower fully-filled spin-up band to the upper empty spin-down band.
-This argument can be verified by our numerical results.
-In fact, the $\delta$ function-like distribution of the spectral weights of a mode in the Stoner continuum [as has be seen in Fig.
+Consequently, an individual spin-1 excitation, i.e., a mode in the Stoner continuum, corresponds to the excitation of an electron from the fully-filled spin-up band to the empty spin-down band with an energy proportional to the difference in the initial and final states plus the band splitting $U/2$.
+Indeed, the $\delta$ function-like distribution of the spectral weights of a mode in the Stoner continuum [as has be seen in Fig.
 \ref{fmcispectrum}($\text{b}_3$) and Fig.
-\ref{fmcispectrum}($\text{b}_4)$] implies that the excitation energy of such a mode can be approximated by the corresponding diagonal term in Eq.
+\ref{fmcispectrum}($\text{b}_4)$] implies that its excitation energy can be approximated by the corresponding diagonal term in Eq.
 (\ref{PHP}).
-By numerics, we also find that the term $\frac{U}{N} \sum_{a=1,2} \sum_{\mathbf{p}\neq\mathbf{k}_i} \left| \mu_{a,\mathbf{p}\uparrow} \right|^2 \left| \mu_{a,\mathbf{k}_{i}-\mathbf{q}\downarrow} \right|^2 \sim \frac{U}{2}$.
-Therefore, the excitation energies of the Stoner continuum are roughly $\varepsilon_{d\downarrow} (\mathbf{k}_i-\mathbf{q}) + \frac{U}{2} -\varepsilon_{d\uparrow} (\mathbf{k}_i)$, which is consistent with the above picture.
-In Fig.
-\ref{fmcispectrum}(a), the lower and upper boundaries of the Stoner continuum approximated by this argument are plotted as the blue solid lines, which fits extremely well with the true boundaries of the shaded area, thus verifies the validity of this simple argument.
-Now we can find out the dominant scattering channels that contribute most to the local boundary bottoms of the Stoner continuum and read off their center-of-mass momentums easily.
-In Fig.
+Numerically, we also find that the term $\frac{U}{N} \sum_{a=1,2} \sum_{\mathbf{p}\neq\mathbf{k}_i} \left| \mu_{a,\mathbf{p}\uparrow} \right|^2 \left| \mu_{a,\mathbf{k}_{i}-\mathbf{q}\downarrow} \right|^2 \sim \frac{U}{2}$.
+Therefore, the excitation energies of the Stoner continuum are roughly $\varepsilon_{d\downarrow} (\mathbf{k}_i-\mathbf{q}) + \frac{U}{2} -\varepsilon_{d\uparrow} (\mathbf{k}_i)$, which is consistent with the above analysis.
+The lower and upper boundaries of the Stoner continuum determined in this way are plotted as the blue solid lines in Fig.
+\ref{fmcispectrum}(a), which fits extremely well with the true boundaries of the shaded area, thus verifies the validity of this simple argument.
+
+\par With this picture, the position of a local boundary bottom of the Stoner continuum is determined by the transferred momentum of the particle-hole scattering between a spin-up maximum and a spin-down minimum.
+Thus, in Fig.
 \ref{bspicture}(b), we mark the local maxima of the spin-up band with red solid circles and the local minima of the spin-down band with purple ones.
-The scattering channels corresponding to the local bottoms marked by the black arrows in Fig.
-\ref{fmcispectrum}(a) are also indicated by green arrows.
-Obviously, a spin-1 excitation corresponding to the particle-hole scattering between a spin-up maximum and a spin-down minimum appears as a local boundary bottom of the Stoner continuum and the momentum difference of these two electron state provides the center-of-mass momentum of that spin-1 excitation.
-To understand the renormalization of the collective modes observed here, in Fig.
-\ref{bspicture}(c), we plot the spectral weights of the four lowest levels subtracted by those in the flatband limit with the center-of-mass momentum $q=(\pi,0)$ ($X_1$ point).
+One can see that the transferred momenta coincide completely with the positions
+ of the local bottoms in Fig. \ref{fmcispectrum}(a) which are indicated here by green arrows.
+
+\par To understand the renormalization of the spin waves observed above, in Fig.
+\ref{bspicture}(c), we plot the spectral weights of the four lowest levels subtracted by those in the flatband limit at momentum $q=(\pi,0)$ ($X$ point).
 The differences concentrate around the scattering channels labeled by $\mathbf{k}_i=(\pm\pi,0)$ and $\mathbf{k}_i=(0,\pm\pi)$.
@@
 Black arrow marks the minimum of the Stoner continuum.
-(a$_2$) and (b$_2$): Low-energy parts of the spin-1 excitation spectra shown in (a$_1$) and (b$_1$) along $\Gamma$-$X_1$ path with an amplified resolution.
+(a$_2$) and (b$_2$): Low-energy parts of the spin-1 excitation spectra shown in (a$_1$) and (b$_1$) along the $\Gamma$-$X$ path with an amplified resolution.
 (a$_3$) and (b$_3$): Illustration of the splitting of the corresponding lower electron bands.
@@
 In Fig.
-\ref{pbcispectrum}($\text{a}_1$) and \ref{pbcispectrum}($\text{b}_1$), we plot the spin-1 excitation spectra of the model near the FM/NFM phase boundary, with the parameters marked by the blue stars in Fig.
+\ref{pbcispectrum}($\text{a}_1$) and \ref{pbcispectrum}($\text{b}_1$), we plot the spin-1 excitation spectra of the model near the FM/NFM phase boundary, with two sets of parameters marked by the blue stars in Fig.
 \ref{phase}(a).
-The corresponding illustrations of the split-band picture are shown in Fig.
+The corresponding illustrations of the band-splitting picture are shown in Fig.
 \ref{pbcispectrum}($\text{a}_3$) and \ref{pbcispectrum}($\text{b}_3$).
 Combined with Fig.
-\ref{fmcispectrum}(a), it can be seen that increased nonflatness leads the moving of the boundary bottom marked by $q_1$ toward zero energy.
-Furthermore, roton-like local minima on the acoustic band appears nearby, which can be seen more clearly in Fig.
+\ref{fmcispectrum}(a), it can be seen that the increase of the nonflatness leads the boundary bottom marked by $q_1$ to move toward low energies further.
+In this process, a roton-like local minimum in the acoustic band appears nearby, which can be seen more clearly in Fig.
 \ref{pbcispectrum}($\text{a}_2$) and \ref{pbcispectrum}($\text{b}_2$).
-With the increase of the nonflatness of the electron band, the energy of this newly formed roton mode goes down and touches zero, which leads to the destabilization of the ground state.
+With the increase of the nonflatness of the electron band, the energy of this newly formed roton-like mode goes down and touches zero, which leads to the destabilization of the ground state.
 From Fig.
@@
 \includegraphics[width=\columnwidth]{tispectrum}
-\caption{(Color online) (a$_1$) and (b$_1$): Spin-1 excitation spectra of the $Z_2$ Hubbard model.
+\caption{(Color online) (a1) and (b1): Spin-1 excitation spectra of the $Z_2$ Hubbard model.
 Green solid lines denote the spin waves and grey regions denote the Stoner continuum.
@@
 \ref{phase}(b) are plotted in Fig.
-\ref{tispectrum}($\text{a}_1$) and Fig.
-\ref{tispectrum}($\text{b}_1$).
-As is similar to Sec.
-\ref{CHM}, the low-lying green solid lines represent the spin waves (collective modes) and the high-energy shaded areas represent the Stoner continuum (individual modes).
-One distinction from the Chern Hubbard model is that the collective modes here are gapped because the $Z_2$ Hubbard model explicitly breaks the spin $SU(2)$ rotation symmetry and no spontaneous continuous symmetry breaking occurs in the ferromagnetic phase.
-The corresponding spectra in the flatband limit are also shown in Fig.
+\ref{tispectrum}($\text{a}_1$) ($t_2=0.650, U=2.0$) and Fig.
+\ref{tispectrum}($\text{b}_1$) ($t_2=0.741, U=2.0$), respectively.
+Same as the case of the Chern insulator discussed in Sec.
+\ref{CHM}, the low-lying green solid lines are identified as the spin waves (collective modes) and the high-energy shaded areas the Stoner continuum (individual modes).
+Overall, the whole spectra are different from those in the Chern insulator shown in Fig.
+\ref{fmcispectrum}(a).
+In particular, the collective modes here are gapped because the $Z_2$ Hubbard model explicitly breaks the spin $SU(2)$ rotation symmetry and no spontaneous continuous symmetry breaking occurs in the ferromagnetic phase.
+Besides, the acoustic and optical bands are degenerate along the $\Gamma$-$X$ path, instead of the $X$-$M$ path reported in the subsection \ref{CHM}.
+
+\par The corresponding spectra in the flatband limit are also shown in Fig.
 \ref{tispectrum}($\text{a}_1$) and Fig.
@@
 \cite{DG_PRB2015}].
-On the one hand, the band bottom of the spin waves in our results lies at the $M_{1,2,3,4}$ points while theirs lies at the $\Gamma$ point.
-On the other hand, the acoustic branch and optical branch are degenerate along the $\Gamma$-$X$ path in our results while they are degenerate along the $X$-$M$ path in theirs.
-These discrepancies can be attributed to the self-interactions of the magnons which are ignored in their scheme but are treated exactly in our method.
-The above observation implies that to get correct dispersion relations, self-interactions between magnons should not be neglected for the $Z_2$ Hubbard model.
+On the one hand, the band bottom of the spin waves in our results lies at the $M$ points while theirs lies at the $\Gamma$ point.
+On the other hand, the acoustic and optical branches are degenerate along the $\Gamma$-$X$ path, which is same as the results for a dispersive band as discussed in the above.
+This is in sharp contrast to that obtained in Ref.
+\cite{DG_PRB2015}, which are degenerate along the $X$-$M$ path.
+These discrepancies are attributed to the effects of multimagnon processes which are ignored in their scheme.
+The above observation suggests that the interactions between magnons have an essential effect on the spin excitations in the $Z_2$ Hubbard model, though they have a negligible effect in the Chern Hubbard model as discussed above.
 
-\par Similar to the Chern Hubbard model, the nonflatness of the $d$ band introduces dips of the Stoner continuum, and the energies of the spin waves around the boundary bottoms of the Stoner continuum are renormalized downward, which results in roton-like modes on both branches of the spin waves.
-The dominant scattering channels leading to the dips of the Stoner continuum and the renormalization of the spin waves can also be determined by the previously proposed band-splitting picture, which are illustrated in Fig.
+\par Similar to the Chern Hubbard model, the introduction of the dispersion to the flatband leads to the appearance of three local minima (roton-like modes) along the $\Gamma-X-M-\Gamma$ direction for both acoustic and optical bands, respectively.
+The momentum positions of these minima coincide the corresponding boundary bottoms of the Stoner continuum, suggesting that the renormalization of the spin-wave bands is due to the coupling between the spin waves and Stoner continuum.
+In the band-splitting picture as proposed above, the position of a local boundary bottom of the Stoner continuum is determined by the transfer momentum of the particle-hole scattering between a spin-up maximum and a spin-down minimum.
+Therefore, we present in Fig.
+\ref{tispectrum}($\text{a}_2$) and \ref{tispectrum}($\text{b}_2$) the dispersions for the lower electron bands in the $Z_2$ Hubbard model showing the splitting of the spin-up and spin-down bands.
+The corresponding local maxima (red solid circles) of the spin-up bands and local minima (purple solid circles) of the spin-down bands in the first Brillouin zone are shown in Fig.
+\ref{tispectrum}($\text{a}_3$) and \ref{tispectrum}($\text{b}_3$).
+One can see that the momentum positions of the boundary bottoms as marked by the black arrows in Fig.
+\ref{tispectrum}($\text{a}_1$) and Fig.
+\ref{tispectrum}($\text{b}_1$) (here only $q_1$ is plotted for illustration) match the transferred momenta between the spin-up and spin-down bands, which gives a firm support of the picture based on the band splitting.
+
+\par From Fig.
+\ref{tispectrum}($\text{a}_1$) to Fig.
+\ref{tispectrum}($\text{b}_1$), the ratio of the Hubbard interaction to the lower electron bandwidth $U/W$ decreases.
+In this process, the roton-like mode with momentum $q_1$ approaches to zero, indicating the destabilization of the ferromagnetic ground state.
+So, the parameters used to get the result in Fig.
+\ref{tispectrum}($\text{b}_1$) marks the boundary between a ferromagnetism phase and a nonferromagnetic phase as shown in the phase diagram Fig.
+\ref{phase}(b).
+From the band splittings shown in Fig.
 \ref{tispectrum}($\text{a}_2$) and Fig.
+\ref{tispectrum}($\text{b}_2$), one can see that the relative energies at the $\Gamma$ and $M$ points increases while that at the $X$ ($Y$) points decreases in this process.
+Consequently, although the momentum positions of the forementioned boundary bottoms remain the same, the corresponding scattering channels are in fact different, which are shifted from $(0,0)\rightarrow(\frac{\pi}{2},\frac{\pi}{2})$ and $(-\pi,-\pi)\rightarrow(-\frac{\pi}{2},-\frac{\pi}{2})$ to $(-\pi,0)\rightarrow(-\frac{\pi}{2},\frac{\pi}{2})$ and $(0,-\pi)\rightarrow(\frac{\pi}{2},-\frac{\pi}{2})$, as shown in Fig.
+\ref{tispectrum}($\text{a}_3$) and Fig.
+\ref{tispectrum}($\text{b}_3$).
+In fact, we find that the energy of the spin-up band at $O$ almost overlaps with the bottom of the spin-down band at $X$ as shown in Fig.
 \ref{tispectrum}($\text{b}_2$).
-Figure \ref{tispectrum}($\text{a}_3$) and Figure \ref{tispectrum}($\text{b}_3$) show the corresponding local maxima (red solid circles) of the spin-up bands and local minima (purple solid circles) of the spin-down bands in the first Brillouin zone.
-Here, only the scattering channels that contribute to the boundary bottoms marked by the black arrows in Fig.
-\ref{tispectrum}($\text{a}_1$) and Fig.
-\ref{tispectrum}($\text{b}_1$) are presented as their induced roton-like modes lead to the destabilization of the ferromagnetic ground state [See Fig.
-\ref{tispectrum}($\text{b}_1$)].
-We want to remark that although the center-of-mass momentums of the forementioned boundary bottoms are the same, the scattering channels are different because the bottom of the spin-down band changes from the $\Gamma$/$M_{1,2,3,4}$ points to the $X_{1,2}$/$Y_{1,2}$ points when $t_2$ changes from $0.650$ to $0.741$ [See Fig.
-\ref{tispectrum}($\text{a}_2$) and Fig.
-\ref{tispectrum}($\text{b}_2$)].
-Last but not least, the ferromagnetic ground state is also quite robust against the spin flips until the bottom of the spin-down band almost overlaps with the top of the spin-up band on energy, which means the phase boundary of the FM/NFM is also quite close to $U/W=2$.
+This result demonstrates that the ferromagnetic ground state is also quite robust against spin flips until the indirect gap between the spin-down and spin-up bands vanishes, which means the phase boundary of the FM/NFM is also quite close to $U/W=2$ from the discussion in Sec.
+\ref{CHM}.
 
@@
 For the Chern Hubbard model, the spin wave is gapless while for the $Z_2$ Hubbard model, the spin wave is gapped.
-Remarkably, for both cases, the nonflatness of the free electron bands introduces dips of the lower boundary of the Stoner continuum, and significantly renormalizes the energies of the collective modes around these dips downward, which can lead to roton-like spin wave excitations.
-With the increase of this nonflatness, the energy of the induced roton-like modes goes down and finally touches zero, which results in the destabilization of the ferromagnetic phase.
+Remarkably, in both cases, the nonflatness of the free electron bands introduces dips in the lower boundary of the Stoner continuum.
+As a result, it renormalizes significantly the energies of the collective modes around these dips through the couplings between spin waves and individual modes in the Stoner continuum, and leads to roton-like spin wave excitations.
+We find that the destabilization of the ferromagnetic phase arises from the softening of the roton-like mode, whose energy goes down gradually with the increase of this nonflatness.
 
-\par The collective modes of the spin-1 excitation spectra in the gap of the Stoner continuum can be viewed as bound states of the individual spin-1 excitations.
-Therefore the physics of the downward renormalization of the energies of the collective modes induced by the boundary bottoms of the Stoner continuum may be quite similar to that observed in the two-dimensional antiferromagents, if we interpret spin waves as bound states of confined spinons and attribute the suppression of the energies of the spin waves to their couplings to the nearby deconfined spinon continuum there\cite{ZFSMC_PRL2006,TS_PRL2013,DMCNTPEMIR_NP2015,SQCCMS_PRX2017,YWDYL_arXiv2018}.
-Although correlated nearly-flat topological bands are hard to realize in real materials, they are possible to be designed in cold-atom systems.
+\par We would like to remark that the physics of the downward renormalization of the energies of the collective modes affected by the boundary bottoms of the Stoner continuum shares similarity with that observed in the two-dimensional antiferromagents, where the spin waves are interpreted as bound states of confined spinons and the suppression of energies of spin waves around the continuum bottom is attributed to their couplings to the nearby deconfined spinon continuum\cite{ZFSMC_PRL2006,TS_PRL2013,DMCNTPEMIR_NP2015,SQCCMS_PRX2017,YWDYL_PRB2018}.
+In addition, although correlated nearly-flat topological bands are hard to realize in real materials, they are possible to be designed in cold-atom systems.
 In fact great progress has been made very recently \cite{LCJPS_N2009,AALBPB_PRL2013,MSKBK_PRL2013,JMDLUGE_N2014}.
-We thus expect our results to be tested in future cold-atom experiments.
+We expect our results will stimulate related investigations in future cold-atom experiments.
 

### 2018-10-30 paper/figure.py

@@ -54,21 +54,12 @@ def lattice():
     p0,p1,p2,p3,p4=np.array([0.0,0.0]),np.array([-1.0,-1.0]),np.array([-1.0,1.0]),np.array([1.0,1.0]),np.array([1.0,-1.0])
     for sp,ep in [(p1,p2),(p2,p3),(p3,p4),(p4,p1),(p0,p3)]:
         ax.plot([sp[0],ep[0]],[sp[1],ep[1]],lw=2,color='black',zorder=1)
-    for x,y in ([0.5,0.5],[-0.5,0.5],[-0.5,-0.5],[0.5,-0.5]):
-        ax.scatter(x,y,s=50,edgecolors='none',color='black',alpha=1.0,marker='o',zorder=3)
-    ax.text(-0.03,+0.03,'$\Gamma$',fontsize=18,va='bottom',ha='right',color='black')
-    ax.text(+0.97,-0.03,'$X_1$',fontsize=18,va='top',ha='right',color='black')
-    ax.text(-1.03,-0.03,'$X_2$',fontsize=18,va='top',ha='right',color='black')
-    ax.text(-0.05,+1.03,'$Y_1$',fontsize=18,va='bottom',ha='right',color='black')
-    ax.text(-0.05,-1.03,'$Y_2$',fontsize=18,va='top',ha='right',color='black')
-    ax.text(+1.00,+1.03,'$M_1$',fontsize=18,va='bottom',ha='center',color='black')
-    ax.text(-1.00,+1.03,'$M_2$',fontsize=18,va='bottom',ha='center',color='black')
-    ax.text(-1.00,-1.03,'$M_3$',fontsize=18,va='top',ha='center',color='black')
-    ax.text(+1.00,-1.03,'$M_4$',fontsize=18,va='top',ha='center',color='black')
-    ax.text(+0.60,+0.45,'$O_1$',fontsize=18,va='top',ha='center',color='black')
-    ax.text(-0.55,+0.45,'$O_2$',fontsize=18,va='top',ha='center',color='black')
-    ax.text(-0.50,-0.55,'$O_3$',fontsize=18,va='top',ha='center',color='black')
-    ax.text(+0.60,-0.55,'$O_4$',fontsize=18,va='top',ha='center',color='black')
+    ax.scatter(0.5,0.5,s=50,edgecolors='none',color='black',alpha=1.0,marker='o',zorder=3)
[…]

### 2018-11-08 paper/paper.tex
For Submission to PRB.

@@
 \author{Zhao-Yang Dong}
+\affiliation{Department of Applied Physics, Nanjing University of Science and Technology, Nanjing 210094, China.}
 \affiliation{National Laboratory of Solid State Microstructures and Department of Physics, Nanjing University, Nanjing 210093, China}
@@
 
-\par Another involved intriguing phenomenon arising from Coulomb repulsions in flat or nearly-flat bands is the itinerant ferromagnetism \cite{T_PTP1998}.
+\par Another involved intriguing phenomenon arising from Coulomb repulsions in flat or nearly-flat bands is the itinerant ferromagnetism \cite{T_PTP1998,T_PRL1992,M_PLA1993,MT_CMP1993,KMTT_EPL2010}.
 It was proved by Tasaki and Mielke that the ground state of a flat electron band with a filling factor not more than but sufficiently close to $1/2$ is ferromagnetically ordered as long as an infinitesimal onsite Hubbard repulsion is present \cite{T_PRL1992,M_PLA1993,MT_CMP1993}.
@@
 \subsection{Exact diagonalization with projection}\label{exact_diag_proj}
-\par Exact diagonalization method with a projection onto the low-energy Hilbert space has been widely applied to systems that host flat or nearly-flat energy bands \cite{NSCM_PRL2011,RB_PRX2011,NSRCM_PRB2011,SGDL_PRB2018,NSRCM_PRL2012,DG_PRB2015}.
+\par Exact diagonalization method with a projection onto the low-energy Hilbert space has been widely applied to systems that host flat or nearly-flat energy bands \cite{NSCM_PRL2011,RB_PRX2011,NSRCM_PRB2011,SGDL_PRB2018,NSRCM_PRL2012}.
 This approach applies when the energy gap between the flat or nearly-flat band and other bands is larger than the Coulomb interaction.
@@
 As discussed in Sec.
-\ref{exact_diag_proj}, the phase boundary is determined as the onset point of $U$ below which negative spin-1 excitation spectra arise.
+\ref{exact_diag_proj}, the phase boundary is determined by the onset parameter at which the spin-1 excitation spectrum starts to acquire zero energy at a finite momentum.
 The phase diagrams of these two models appear quite similar.
@@
 \ref{fmcispectrum}(a).
+This observation shows that it is the extending of the Stoner continuum arising from the nonflatness of the $d$ band that pushes down the spin-wave excitations.
+In particular, it suggests that the couplings between the individual excitations and spin waves would be responsible for the emergence of the roton-like spin waves.
 
@@
 \end{figure}
-
-\par This observation shows that it is the extending of the Stoner continuum arising from the nonflatness of the $d$ band that pushes down the spin-wave excitations.
-In particular, it suggests that the couplings between the individual excitations and spin waves would be responsible for the emergence of the roton-like spin waves.
 
@@
 \ref{pbcispectrum}($\text{a}_3$) and \ref{pbcispectrum}($\text{b}_3$).
-Combined with Fig.
+Compared with Fig.
 \ref{fmcispectrum}(a), it can be seen that the increase of the nonflatness leads the boundary bottom marked by $q_1$ to move toward low energies further.
@@
 \includegraphics[width=\columnwidth]{tispectrum}
-\caption{(Color online) (a1) and (b1): Spin-1 excitation spectra of the $Z_2$ Hubbard model.
+\caption{(Color online) (a$_1$) and (b$_1$): Spin-1 excitation spectra of the $Z_2$ Hubbard model.
 Green solid lines denote the spin waves and grey regions denote the Stoner continuum.
@@
 The momentum positions of these minima coincide the corresponding boundary bottoms of the Stoner continuum, suggesting that the renormalization of the spin-wave bands is due to the coupling between the spin waves and Stoner continuum.
-In the band-splitting picture as proposed above, the position of a local boundary bottom of the Stoner continuum is determined by the transfer momentum of the particle-hole scattering between a spin-up maximum and a spin-down minimum.
+In the band-splitting picture as proposed above, the position of a local boundary bottom of the Stoner continuum is determined by the transfer momentum of the particle-hole scattering between a spin-up band maximum and a spin-down band minimum.
 Therefore, we present in Fig.
@@
 \ref{tispectrum}($\text{a}_1$) and Fig.
-\ref{tispectrum}($\text{b}_1$) (here only $q_1$ is plotted for illustration) match the transferred momenta between the spin-up and spin-down bands, which gives a firm support of the picture based on the band splitting.
+\ref{tispectrum}($\text{b}_1$) (here only $q_1$ is plotted for illustration) match the transferred momenta between the spin-up and spin-down bands, which gives a strong support of the picture based on the band splitting.
 
@@
 
-\par We would like to remark that the physics of the downward renormalization of the energies of the collective modes affected by the boundary bottoms of the Stoner continuum shares similarity with that observed in the two-dimensional antiferromagents, where the spin waves are interpreted as bound states of confined spinons and the suppression of energies of spin waves around the continuum bottom is attributed to their couplings to the nearby deconfined spinon continuum\cite{ZFSMC_PRL2006,TS_PRL2013,DMCNTPEMIR_NP2015,SQCCMS_PRX2017,YWDYL_PRB2018}.
+\par We would like to remark that the downward renormalization of the energies of the collective modes affected by the boundary bottoms of the Stoner continuum shares similarity with that observed in the two-dimensional antiferromagents, where the spin waves are interpreted as bound states of confined spinons and the suppression of energies of spin waves around the continuum bottom is attributed to their couplings to the nearby deconfined spinon continuum\cite{ZFSMC_PRL2006,TS_PRL2013,DMCNTPEMIR_NP2015,SQCCMS_PRX2017,YWDYL_PRB2018}.
 In addition, although correlated nearly-flat topological bands are hard to realize in real materials, they are possible to be designed in cold-atom systems.
@@
 \begin{acknowledgments}
-\par This work was supported by the National Natural Science Foundation of China (11774152) and National Key Projects for Research and Development of China (Grant No. 2016YFA0300401).
+\par This work was supported by the National Natural Science Foundation of China (No.11774152, No.11674158) and National Key Projects for Research and Development of China (Grant No. 2016YFA0300401).
 \par X.-F. S. and Z.-L. G. contributed equally to this work.

### 2018-12-24 paper/paper.tex
Revised manuscript for PRB.

@@
 \par We would like to remark that the downward renormalization of the energies of the collective modes affected by the boundary bottoms of the Stoner continuum shares similarity with that observed in the two-dimensional antiferromagents, where the spin waves are interpreted as bound states of confined spinons and the suppression of energies of spin waves around the continuum bottom is attributed to their couplings to the nearby deconfined spinon continuum\cite{ZFSMC_PRL2006,TS_PRL2013,DMCNTPEMIR_NP2015,SQCCMS_PRX2017,YWDYL_PRB2018}.
+Intriguing flatband physics arises not only in models with repulsive interactions, but also in those with attractive ones\cite{PT_NC2015,JPVKT_PRL2016,LVPSHT_PRB2017,TLP_arXiv2018}, where the Bardeen-Cooper-Schrieffer superconducting state is the exact ground state in the flatband limit\cite{PT_NC2015,JPVKT_PRL2016}.
+We hope our methods can be generalized to these systems beyond the flatband limit.
 In addition, although correlated nearly-flat topological bands are hard to realize in real materials, they are possible to be designed in cold-atom systems.
